---
title: "A quick introduction to FLR"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
tags: [FLR FLCore]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE}
library(knitr)
source("R/ini.R")
```

The Fisheries Library in R (FLR) is a collection of tools for quantitative fisheries science, developed in the R language, that facilitates the construction of bio-economic simulation models of fisheries systems as well as the application of a wide range of quantitative analysis.

FLR builds on the powerful R environment and syntax to create a [domain-specific language](https://en.wikipedia.org/wiki/Domain-specific_language) for the quantitative analysis of the expected risks and impacts of fisheries management decisions. The classes and methods in FLR consider uncertainty an integral part of our knowledge of fisheries system. [...]

## Required packages

To follow this tutorial you should have installed the following packages:

- FLR: [FLCore](http://www.flr-project.org/FLCore/)

You can do so as follows:

```{r, eval=FALSE}
install.packages(c("FLCore"), repos="http://flr-project.org/R")
```

# Getting started with the FLCore classes

The main *classes* (i.e. data structures) and *methods* (i.e. procedures) in the FLR system are found in the `FLCore` package. Let's load it first.

```{r, pkgs}
library(FLCore)
```

We can then inspect an example object.

```{r, ple4}
data(ple4)
```

The `ple4` object is of class `FLStock`, used in **FLR** to represent a fish population that is constructed from catch and abundance data through a stock assesssment model. `FLStock` is an S4 class (see `?Classes_Details` for futher details), consisting of a number of slots designed to each of the elements, data or results, needed to construct a view of a fish stock. By calling the `summary` method on the object

```{r, summaryple4}
summary(ple4)
```

we can inspect the slots, dimensions and structure. Most slots in the class (e.g. `catch` or `stock.n`) are themselves of another **FLCore** class, `FLQuant`. This class, the basic element used to assemble all other classes in **FLR**, is a 6-dimensional array that can take advantage of the powerful array algebra capabilities of R. All slots can be accessed and modified using accessors and replacement methods.

```{r, accessors}
catch(ple4)
m(ple4) <- m(ple4) + m(ple4) * 0.5
```

Other standard R methods have also been defined for these classes in a way that is as intuitive as possible for any R user. For example, subsetting using the `[` operator works on both `FLStock`

```{r, subsetFLStock}
summary(ple4[, 1:10])
```

and `FLQuant`

```{r, subsetFLQuant}
stock.n(ple4)[1,]
```

while ensuring that the result are always valid objects of the same class. There are, however, necessary difference. For example, selecting a single element along the first dimension (`age`) did not drop that dimension from the object (`drop=FALSE`), in contrast with the standard behaviour in R for `array`.

In addition to the `summary` method above, a creator method is available for each class

```{r, creator}
FLQuant(rlnorm(30), dimnames=list(age=0:5, year=2012:2017))
```

Other common methods have also been defined for all of them, for example to be able to coerce to and from other classes,

```{r, as.df}
head(as.data.frame(ple4))
```

or plot an object.

```{r, plotFLStock, fig.cap="FLStock plot for ple4"}
plot(ple4)
```

A number of fisheries-specific calculations are also available. For example, the estimated spawning stock biomass ($SSB$), can be obtained from an `FLStock` object using

```{r, ssb}
ssb(ple4)
```

while the mean fishing mortality accross the fully-selected ages ($\bar{F}$) is calculated with

```{r, fbar}
fbar(ple4)
```

## Class validity

The S4 classes defined in **FLCore**  all have validity functions defined that limit what changes can be made to an object for it to remain valid. This ensures that methods do not encounter objects that do not have the required dimensions, differ in dimension names, or are not compatible with each other. For example, the validity requirements for the `FLQuant` class require that

- it is a 6-dimensional array,
- the array is numeric,
- the first dimension is not named 'cohort', and
- dimensions 2:5 are named 'year, 'unit' 'season', 'area' and 'iter'.

Object validity is checked not only by the class constructor, but also by the replacement methods, for example when calling

```{r, validity}
catch(ple4) <- landings(ple4) + discards(ple4)
```

to assign a new value (the result of adding `landings` and `discards`) to a single slot (`catch`).

# A modelling example: the FLSR class

A different type of class defined in **FLCore** is `FLSR`. This class allows for fitting, exploration and prediction of a stock-recruitment relationship. An example dataset is also available, corresponding to the North Sea herring stock.

```{r, nsher}
data(nsher)
summary(nsher)
```

The class holds together `FLQuant` slots for inputs (`rec` for recruitment in numbers and `ssb` for spawning stock biomass or any other proxy of stock reproductive potential) and outputs of the fit (`params` for the estimated parameters, `fitted` for the estimated recruitment, and `residuals` for the log residuals). It also contains the necessary elements for the model fit to be carried out using maximum likelihood estimation:

- `logl`: a function that computes the log-likehood from observations and parameter values, to be passed on to `optim`,
- `model`: a formula to calculate the estimated recruitment, and
- `initial`: a function to obtain initial parameter values for the minimization algorithm.

By calling the method that carries out the minimization, `fmle`, we obtain a new object in which the results of the fit are now available:

- `params`: the estimated parameters,
- `logLik`: the log-likelihood and degrees of freedom,
- `covar`: the variance-covariance matrix of the fit, and
- `hessian`: the Hessian matrix of the fit,

as well as some other information returned by the fitting procedure.

Of course we can visualize the result of the model fit, together with a useful set of diagnostics, by simply calling the `plot` method for the class

```{r, plotFLSR}
plot(nsher)
```

# FLBiol

# Packages

# References

L. T. Kell, I. Mosqueira, P. Grosjean, J-M. Fromentin, D. Garcia, R. Hillary, E. Jardim, S. Mardle, M. A. Pastoors, J. J. Poos, F. Scott, R. D. Scott; FLR: an open-source framework for the evaluation and development of management strategies. *ICES J Mar Sci* 2007; 64 (4): 640-646. doi: [10.1093/icesjms/fsm012](https://doi.org/10.1093/icesjms/fsm012).

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Alternatively, send a pull request to <https://github.com/flr/doc/>.
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage: <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission, DG Joint Research Centre, Directorate D - Sustainable Resources, Unit D.02 Water and Marine Resources, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>.
