---
title: Reading data into FLR 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---

```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE}
library(knitr)
source("R/ini.R")
```
 
This tutorial details methods for reading data in various formats into R for generating objects of the [FLStock](http://www.flr-project.org/FLCore/FLStock.html), [FLIndex](http://www.flr-project.org/FLCore/FLStock.html) and [FLFleet](http://www.flr-project.org/FLCore/FLStock.html) classes.

## Required packages

To follow this tutorial you should have installed the following packages:

- CRAN: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/); [FLFleet](http://www.flr-project.org/FLFleet/); [ggplotFL](http://www.flr-project.org/ggplotFL/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("ggplot2"))
install.packages(c("FLCore","ggplotFL"), repos="http://flr-project.org/R")
```

```{r, pkgs}
# Load all necessary packages, trim pkg messages
library(FLFleet)
library(ggplotFL)
```

## Example data files

The data files used in this tutorial need to be available in R's working directory, and this code obtains them from the **FLR** website. Files will be downloaded to a temporary folder. If you want to keep a local copy, simply set a different value to the `dir` variable below.

```{r, getfiles, message=FALSE}
dir <- tempdir()
download.file("http://flr-project.org/doc/src/loading_data.zip", file.path(dir, "loading_data.zip"))
unzip(file.path(dir, "loading_data.zip"), exdir=dir)
```

### Reading files (csv, dat, ...)

Fisheries data are generally stored in different formats (cvs, excel, SAS...). R provides tools to read and import data, from simple text files to more advanced SAS files or databases. [Datacamp](https://www.datacamp.com/community/tutorials/importing-data-r-part-two#gs.kNzBd5k) is a useful tutorial to quickly import data into R.

Your data are stored in a folder in your computer, or on a server. R requires a path to the data. You can check the working directory already active in your R session using the command `getwd()`. To set the working directory use `setwd("directory name")`. Case is important. Use '//' or '\' for separating folders and directories in Windows.

This tutorial will give some examples, but regardless of the format, the different steps are:

- Finding the right function to import data into R
- Reshaping the data as a matrix 
- creating an FLQuant object

## Importing files into R (example of csv file)

There are many ways of reading csv files. [read.table](http://stat.ethz.ch/R-manual/R-devel/library/utils/html/read.table.html) with 'header', 'sep', 'dec' and 'row.names' options will allow you to read all .csv and .txt files.

The [read.csv](http://stat.ethz.ch/R-manual/R-devel/library/utils/html/read.table.html) or [read.csv2](http://stat.ethz.ch/R-manual/R-devel/library/utils/html/read.table.html) functions are very useful for reading csv files. 

```{r, readcsv}
catch.n <- read.csv(file.path(dir,"catch_numbers.csv"), row=1)
# We have read in the data as a data.frame
class(catch.n)
```

The data are now in your R environment. Before creating an **FLQuant** object, you need to make sure it is consistent with the type of object and formatting that is needed to run the [FLQuant](http://www.flr-project.org/FLCore/FLQuant.html) method. To get information on the structure and format needed, type ?FLQuant in your R Console.

## Reshaping data as a matrix 

FLQuant objects accept 'vector', 'array' or 'matrix'. We can convert the object *catch.n* to a matrix.

``` {r, reshape}
catch.n.matrix <- as.matrix(catch.n)
catch.n.matrix[,1:8]
```

An [FLQuant](http://www.flr-project.org/FLCore/FLQuant.html) object is made of six dimensions. The name of the first dimension can be altered by the user from its default, 'quant'. This could typically be 'age' or 'length' for data related to natural populations. The only name not accepted is 'cohort', because data structured along cohort should be stored using the [FLCohort](http://www.flr-project.org/FLCore/FLCohort.html) class instead. Other dimensions are always named as follows: 'year' for the calendar year of the data point; 'unit' for any kind of division of the population, e.g. by sex; 'season' for any temporal strata shorter than year; 'area' for any kind of spatial stratification; and 'iter' for replicates obtained through bootstrap, simulation or Bayesian analysis.

When importing catch numbers, for example, the input object needs to be formatted as such: age or length in the first dimension and year in the second dimension. If the object is not formatted in the right way, you can use the `reshape` functions from the package [reshape2](https://cran.r-project.org/web/packages/reshape2/index.html).

## Making an FLQuant object 

We need to specify the dimnames

```{r, flquant}
catch.n.flq <- FLQuant(catch.n.matrix, dimnames=list(age=1:7, year = 1957:2011))
catch.n.flq[,1:7]
```

## Reading common fisheries data formats 

FLCore contains functions for reading in fish stock data in commonly used formats. To read a single variable (e.g. numbers-at-age, maturity-at-age) from the **Lowestoft VPA** format you use the `readVPA` function. The following example reads the catch numbers-at-age for herring:

```{r, readVPA}
# Read from a VPA text file
catch.n <- readVPAFile(file.path(dir, "her-irlw","canum.txt"))
class(catch.n)
```

This can be repeated for each of the data files. In addition, functions are available for [Multifan-CL](http://www.multifan-cl.org/) format `readMFCL`, and [ADMB](http://www.admb-project.org/) format `readADMB`.

Alternatively, if you have the full information for a stock in the **Lowestoft VPA**, **Adapt**, **CSA** or **ICA** format, you can read it all in together using the `readFLStock` function. Here, you point the function to the index file, with all other files in the same directory:

```{r, readFLStock}
# Read a collection of VPA files, pointing to the Index file:
# DELETE: her <- readFLStock('http://flr-project.org/doc/src/her-irlw/index.txt')
her <- readFLStock(file.path(dir, 'her-irlw', 'index.txt'))
class(her)
```

This correctly formats the data as an [FLStock](http://www.flr-project.org/FLCore/FLStock.html) object. Note: the units for the slots have not been set. We will deal with this in the next section.

```{r, readFLStock2}
summary(her)
```

This object only contains the input data for the stock assessment, not any estimated values (e.g. harvest rates, stock abundances). You can add these to the object as follows:

```{r, AddMissingAssessmentData}
stock.n(her) <- readVPAFile(file.path(dir, "her-irlw", "n.txt"))
print(stock.n(her)[,ac(2007:2011)]) # only print 2007:2011
harvest(her) <- readVPAFile(file.path(dir,"her-irlw", "f.txt"))
```

Now we have a fully filled [FLStock](http://www.flr-project.org/FLCore/FLStock.html) object. But let's check the data are consistent.

```{r, CheckConsistency}
# The sum of products (SOP)
apply(landings.n(her)*landings.wt(her), 2, sum)[,ac(2007:2011)]
# and the value read in from the VPA file
landings(her)[,ac(2007:2011)]
## They are not the same!!  We correct the landings to be the same as the SOP - there is a handy function for this purpose
landings(her) <- computeLandings(her)
# In addition, there is no discard information
discards.wt(her)[,ac(2005:2011)]
discards.n(her)[,ac(2005:2011)]
# Set up the discards and catches
discards.wt(her)  <- landings.wt(her)
discards.n(her)[] <- 0
discards(her)     <- computeDiscards(her)
catch(her)        <- landings(her)
catch.wt(her)     <- landings.wt(her)
catch.n(her)       <- landings.n(her)
```

Functions are available to [computeLandings](http://www.flr-project.org/FLCore/computeLandings.html), [computeDiscards](http://www.flr-project.org/FLCore/computeDiscards.html), [computeCatch](http://www.flr-project.org/FLCore/computeCatch.html) and [computeStock](http://www.flr-project.org/FLCore/computeStock.html). These functions take the argument slot = 'catch', slot = 'wt' and slot = 'n' to compute the total weight, individual weight and numbers respectively, in addition to slot = 'all'. 

## Adding a description, units, ranges etc..

Before we are finished, we want to ensure the units and range references are correct. This is important as the derived calculations require the correct scaling (e.g. `fbar`, for the average fishing mortality range over the required age ranges).

First, let's ensure an appropriate name and description are assigned:

```{r, Descriptions}
summary(her)
#name and descriptions
name(her) # ok
desc(her) # ok
# Set the Fbar range for the stock 
range(her)[c('minfbar','maxfbar')] # ok, but can be filled with  <- c(min,max)
# set the plus group
range(her)['plusgroup']  <- 7 # final year is a plusgroup
## Units
units(catch(her)) <- units(discards(her)) <- units(landings(her)) <- units(stock(her)) <- 'tonnes'
units(catch.n(her)) <- units(discards.n(her)) <- units(landings.n(her)) <- units(stock.n(her)) <- '1000'
units(catch.wt(her)) <- units(discards.wt(her)) <- units(landings.wt(her)) <- units(stock.wt(her)) <- 'kg'
units(harvest(her)) <- 'f'
```
This should now have the correct units defined:

```{r, Descriptions2}
summary(her)
plot(her) + theme_bw() # using the simple black and white theme
```

## FLIndex objects

Two solutions can be used to read abundance indices into FLR. 

If your data are formatted in a **Lowestoft VPA** format, then [FLCore](http://www.flr-project.org/FLCore/) contains functions for reading in indices. To read an abundance index, you use the `readFLIndices` function. The following example reads the index from the `ple4` example:

```{r, readFLIndices}
indices <- readFLIndices(file.path(dir, "ple4_ISIS.txt"))
```

Using this function, the 'names' and 'range' slots are already filled.

If your data are not formatted in a **Lowestoft VPA** format, then you and read them using [read.table](http://stat.ethz.ch/R-manual/R-devel/library/utils/html/read.table.html) from base R, for example.

```{r, readFLIndices2}
indices <- read.table(file.path(dir, "ple4Index1.txt"))
# transform into FLQuant
indices <- FLQuant(as.matrix(indices), dimnames=list(age=1:8, year = 1985:2008))
# and into FLIndex
indices <- FLIndex(index = indices)
# and then into FLIndices
indices <- FLIndices(indices)
plot(indices[[1]])
```

The 'range' slot needs to be filled with the end and start date of the tuning series

```{r, rangeslot}
range(indices[[1]])[c('startf', 'endf')] <- c(0.66,0.75)
```

## FLFleet objects

Reading data for fleets into an [FLFleet](http://www.flr-project.org/FLFleet/) object is complicated by the multi-layer structure of the object. The object is defined so that:

```{r, leveltable, echo = F}
kable(data.frame(Level = c(1,2,3),
		 Class = c('FLFleet','FLMetier(s)','FLCatch(es)'),
		 Contains = c('variables relating to vessel level activity',
			            'variables relating to fishing level activity',
		              'variables relating to stock catches')))
```

Here are the slots for each level:

```{r, FLFleetslots}
# FLFleet level
summary(FLFleet())
# FLMetier level
summary(FLMetier())
# FLCatch level
summary(FLCatch())
```

Due to the different levels, units and dimensions of the variables, and the potentially high number of combinations of fleets, métiers and stocks in a mixed fishery, getting the full data into an `FLFleets` object (which is a list of [FLFleet](http://www.flr-project.org/FLFleet/) objects) can be an onerous task.

A way of simplifying the generation of the fleet object is to ensure all the data are in a csv file with the following structure:

```{r, Fleetdata, echo = F}
kable(data.frame(Fleet = c('Fleet1', 'Fleet2'),
		             Metier = c('Metier1', 'Metier1'),
		             Stock = c('Stock1', 'Stock2'),
	               type = c('landings.n', 'landings.wt'),
		             age = c(1,1),
		             year = c(2011,2011),
		             unit = c(1,1),
		             season = c('all', 'all'),
		             area = c('unique', 'unique'),
		             iter = c(1,1),
		             data = c(254,0.3)))
```

To generate the required structure, you can then read in the file and generate the object using an [lapply](http://petewerner.blogspot.it/2012/12/using-apply-sapply-lapply-in-r.html) function:

```{r, GeneratingFLFleets, eval = F}
# Example of generating fleets
fl.nam <- unique(data$Fleet) # each of the fleets

yr.range <- 2005:2011 # year range of the data - must be same, even if filled with NAs or 0s

# empty FLQuant for filling with right dimensions
fq  <- FLQuant(dimnames = list(year = yr.range), quant = 'age')

### Fleet level slots ###
fleets <- FLFleet(lapply(fl.nam, function(Fl) {

# blank quants with the same dims
eff <- cap <- crw <- cos.fl <- fq

# fleet effort
eff[,ac(yr.range)] <- data$data[data$Fleet == Fl & data$type == 'effort']
units(eff) <- '000 kw days'

## Repeat for each fleet level variables (not shown) ##

### Metier level slots ###
met.nam  <- unique(data$Metier[data$Fleet == Fl]) # metiers for fleet
met.nam  <- met.nam[!is.na(met.nam)] # exclude the fleet level data

metiers  <- FLMetiers(lapply(met.nam, function(met) {

# blank quants
effmet <- cos.met <- fq

# effort share for metier
effmet[,ac(yr.range)] <- data$data[data$Fleet == Fl & data$Metier & data$type == 'effshare']
units(effmet)  <- NA

## Repeat for each metier level variables (not shown) ##


sp.nam <- unique(data$stock[data$Fleet == Fl & data$Metier == met]) # stocks caught by metier
sp.nam <- sp.nam[!is.na(sp.nam)] # exclude fleet and metier level data

catch <- FLCatches(lapply(sp.nam, function(S){
print(S)

# Quant dims may be specific per stock
la.age <- FLQuant(dimnames = list(age = 1:7, year = yr.range, quant = 'age'))
la.age[,ac(yr.range)] <- data$data[data$Fleet == Fl & data$Metier == met & data$Stock == S & data$type == 'landings.n']
units(la.age) <- '1000'

## Repeat for all stock level variables (not shown) ##

# Build F
res <- FLCatch(range = yr.range, name = S, landings.n = la.age,...)

## Compute any missing slots, e.g.
res@landings <- computeLandings(res)

return(res) # return filled FLCatch

})) # End of FLCatches

# Fill an FLMetier with all the stock catches
m <- FLMetier(catches = catch, name = met)
m@effshare  <- effmet
m@vcost <- vcost

		 })) # end of FLMetiers

fl <- FLFleet(metiers = metiers, name = Fl, effort = ef,...) # fill with all variables
return(fl)

		 }))

names(fleets) <- fl.nam
```

You should now have a multilevel object with `FLFleets` containing a list of [FLFleet](http://www.flr-project.org/FLCore/FLFleet.html) objects, each which in turn contain `FLMetiers` with a list of [FLMetier](http://www.flr-project.org/FLCore/FLMetier.html) objects for the fleet, and a list of `FLCatches` containing [FLCatch](http://www.flr-project.org/FLCore/FLCatch.html) objects for each stock caught by the métier.

# References

None

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR web-page, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* ggplot2: `r packageVersion('ggplot2')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission, DG Joint Research Centre, Directorate D - Sustainable Resources, Unit D.02 Water and Marine Resources, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>


