---
title: Modelling Stock-Recruitment with FLSR
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    github_document
tags:
license: Creative Commons CC-BY SA
---

```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE}
library(knitr)
source("R/ini.R")
```
 
*FLSR* is a [S4](https://stat.ethz.ch/R-manual/R-devel/library/methods/html/Classes_Details.html) class for Stock-Recruitment (SR) models, an extension of *FLModel* , and part of the *FLCore* package.  Commonly used or custom-tailored SR models can be fitted directly on *FLStock* objects and provide estimates of uncertainty. *FLSR* class objects can be then used to visualize the fitted models, in calculations of biological reference points using *FLBPR* or when performing stock projections. 

## Required packages

To follow this tutorial you should have installed the following packages:

- CRAN: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/), [ggplotFL](http://www.flr-project.org/ggplotFL/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("ggplot2"))
install.packages(c("FLCore"), repos="http://flr-project.org/R")
install.packages(c("ggplotFL"), repos="http://flr-project.org/R")
```

Initially, the libraries need to be called.

```{r, pkgs}
# This chunk loads all necessary packages, trims pkg messages
library(FLCore)
library(ggplotFL)
```

And then the user can load and visualize the results of an assessment (VPA) already performed and stored in the ple4 *FLStock* object.

```{r, data}
# This chunk loads the ple4 FLStock object 
data(ple4)
```

```{r figA}
# This chunk plots the assesment output
plot(ple4)
```


# The Stock-Recruitment (SR) relationship

Given that recruitment and spawning stock biomass (SSB) are provided as an output of the assessment, their relationship can be visualized simply by ploting the recruits against the SSB.

```{r figB}
# This chunk plots the SSB-Recruits graph
#plot(FLQuants(ple4, "ssb", "rec"))
ggplot(aes(ssb,rec), data=model.frame(FLQuants(ple4, "ssb", "rec")))+geom_point()+geom_smooth()
```

## Working with FLSR objects
An empty *FLSR* object can be directly created simply by:

```{r, FLSRobject1}
# This chunk creates an empty FLSR object
sr1 <- FLSR()
```

An *FLSR* object can be also be created from converting directly an *FLStock* object:

```{r, FLSRobject2}
# This chunk converts an FLStock object into an FLSR object
p4sr <- as.FLSR(ple4)
```

The contents of the *FLSR* object are the following:

```{r, FLSRobject2_contents}
# This chunk outputs the summary of the FLSR object
summary(p4sr)
```

In the case of ple4 data, recruits are fish of age=1. Hence, the lag between ssb and rec is also 1 year. The starting year for SSB is 1957, whereas for recruits it is 1958. 

```{r, rec_ssb_lag}
# This chunk outputs the contents of the rec and ssb slots of the FLSR object
ssb(p4sr)[,1]
rec(p4sr)[,1]
```

The default recruitment age in FLSR is age=1. However, the user can change this by triming the *FLStock* object while converting it into an *FLSR* object:

```{r, set_rec_age1}
# This chunk shows how to set a different recruitment age than the default, 
# e.g. set the recruitment age at age=2
# this can be done by trimming the FLStock object as follows
p4sr2 <-as.FLSR(ple4[-1])
```

In this case, the lag between ssb and rec is 2 years. The starting year for SSB is 1957, whereas for recruits it is 1959.

```{r, set_rec_age2}
# Note the shift in years, reflecting that recruitment is now at age 2
ssb(p4sr2)[,1]
rec(p4sr2)[,1]
```

## Fitting a SR model 

To fit a SR model a series of commonly-used stock-recruitment models are already available,
including the corresponding likelihood functions and calculation of initial
values. See [SRModels](https://github.com/flr/FLCore/blob/master/R/SRmodels.R) for more details and the exact formulation implemented for each of them. Each method is defined as a function returning a list with one or more elements as follows:

* modelFormula for the model, using the slot names (rec and ssb) to refer to the usual inputs.

* loglFunction to calculate the loglikelihood of the given model when estimated through Maximum Likelihood Estimation (MLE, see <span style="background:yellow">fmle</span>).

* initialFunction to provide initial values for all parameters to be minimization algorithms called by <span style="background:yellow">fmle</span> or <span style="background:yellow">nls</span>. If required, this function also have two attributes, <span style="background:yellow">lower</span> and <span style="background:yellow">upper</span>, that give lower and upper limits for the parameter values, respectively. This is used by some of the methods defined in <span style="background:yellow">optim</span>, like <span style="background:yellow">"L-BFGS-B"</span>. The _model_ <-- method for *FLModel* can then be called with _value_ being a list thus described, the name of the function returning such a list, or the function itself. 

The available SR models are: bevholt(), bevholt.ar1(), bevholt.c.a(), bevholt.c.b(), bevholt.d(), bevholt.ndc(), bevholt.sv(), geomean(), logl.ar1(rho, sigma2, obs, hat), ricker(), ricker.ar1(), ricker.c.a(), ricker.c.b(), ricker.d(), ricker.sv(), segreg(), shepherd(), shepherd.ar1(), shepherd.d(), shepherd.d.ar1(), shepherd.ndc(), shepherd.ndc.ar1(), sv2ab(steepness, vbiomass, spr0, model). 

The user can assign a Ricker SR model to the *FLStock* object. The user can also obtain the model formula of the fitted model, as well as the log-likelihood. The <span style="background:yellow">fmle</span> method fits the model specified in an *FLModel* object using R's <span style="background:yellow">optim</span> function to minimize the negative of the log-likelihood function, in the logl slot, through calls to the minimizaton routine. The default algorithm for optim is Nelder-Mead, however other options are available (e.g. "L-BFGS-B", see ?optim).

```{r, fit_SR_model}
# This chunk assigns a Ricker SR model and fits it
model(p4sr) <- ricker()
model(p4sr)
# the fmle method then fits the SR model using logl and R's optim model fitting through MLE
p4sr<-fmle(p4sr)
# log-likelihood
# logl(p4sr)
```

The user can extract the initial parameters used by the optimiser, as well as the lower and upper limits of these parameters.

```{r, SR_init_params_lmts}
# initial values for the optimiser
initial(p4sr)
# lower and upper limits for the parameters
lower(p4sr)
upper(p4sr)
```

Diagnostic plots can be produced by simply calling the /plot/ function on the *FLSR* object.

```{r, figC}
# Diagnostics plots of the fitted SR model
plot(p4sr)
```

## NS Herring stock-recruitment dataset example

The user can experiment with North Sea herring data where a Ricker model has already been fitted.

```{r, nsher_ricker}
# This chunk loads, plots and prints a summary of the nsher FLSR object 
# (a ricker SR model has already been fitted)
data(nsher)
plot(nsher)
summary(nsher)
```

The user can change the fitted SR model if so desired. Below bevholt() and cushing() models are used.

```{r, nsher_bh_cs}
# This chunk fits and plots a bevholt SR model and a cushing SR model on nsher data
#assign nsher with ricker model to a new object
nsher_ri <- nsher
# change model to bevholt
model(nsher) <- bevholt()
# fit through MLE
nsher_bh <- fmle(nsher)
# change model to cushing
model(nsher) <- cushing()
# fit through MLE
nsher_cs <- fmle(nsher)
```

One can inspect the fits visually,

```{r, ri_bh_cs_plots}
# this chunk plots the fits from the 3 different SR models
plot(nsher_ri)
plot(nsher_bh)
plot(nsher_cs)
```

by using the AIC,

```{r, ri_bh_cs_AIC}
# this chunk plots the fits from the 3 different SR models
print(paste0('Ricker: ',round(AIC(nsher_ri),4),' ',
             'Beverton-Holt: ',round(AIC(nsher_bh),4),' ',
             'Cushing: ',round(AIC(nsher_cs),4)))
```

or Schwarz's Bayesian Information Criterion

```{r, ri_bh_cs_BIC}
# this chunk plots the fits from the 3 different SR models
print(paste0('Ricker: ',round(BIC(nsher_ri),4),' ',
             'Beverton-Holt: ',round(BIC(nsher_bh),4),' ',
             'Cushing: ',round(BIC(nsher_cs),4)))
```

Additionally, a profiling of the model parameters can be visualized for each fitted model.

```{r, figD}
# Profile the likelihood to check the fit
par(mfrow=c(1,3))
profile(nsher_ri)
profile(nsher_bh)
profile(nsher_cs)
```

SR model parameters can also be fixed. In this case, _steepness_ is fixed to a value of 0.8. Details on the model parameterization can be found in  [SRmodels](https://github.com/flr/FLCore/blob/master/R/SRmodels.R).

```{r, figE}
# Fit a bevholtSV model with fixed steepness at 0.8
par(mfrow=c(1,1))
model(p4sr) <- bevholtSV
p4sr <- fmle(p4sr, fixed = list(s = 0.8))
plot(p4sr)
params(p4sr)
```

Custom SR models can be implemented. To define a new model requires the specification of it´s i) functional form, ii) likelihood iii) bounds and iv) starting values. For example, the user can fit the Deriso-Schnute model below.

```{r, SR_custom}
# Fit a custom SR model (Deriso Schnute)

dersch<-function(){
  logl <- function(a,b,c,rec,ssb) {
          res<-loglAR1(log(rec), log(a*ssb*(1-b*c*ssb)^(1/c)))
          return(res)
          }

  ## initial parameter values
  initial <- structure(function(rec, ssb){
     slopeAt0 <- max(quantile(c(rec)/c(ssb), 0.9, na.rm = TRUE))
     maxRec   <- max(quantile(c(rec), 0.75, na.rm = TRUE))

     ## Bevholt by default c=-1
     return(FLPar(a=slopeAt0, b=1/maxRec, c=-1))},

  lower=rep(-Inf, 3),
	upper=rep( Inf, 3))

  model  <- rec~a*ssb*(1-b*c*ssb)^(1/c)

  return(list(logl = logl, model = model, initial = initial))}

model(nsher)<-dersch()
nsher_dersch<-fmle(nsher,fixed=list(c=-1))
plot(nsher_dersch)
```

A SR model with AR1 autocorrelation can be also be fitted. 
```{r, SR_custom_AR1}
# Fit a custom SR AR1 model
rickerAR1 <- function()
  {
  ## log likelihood, assuming normal log.
  logl <- function(a, b, rho, rec, ssb)
      loglAR1(log(rec), log(a*ssb*exp(-b*ssb)), rho=rho)

  ## initial parameter values
  initial <- structure(function(rec, ssb) {
		# The function to provide initial values
    res  <-coefficients(lm(c(log(rec/ssb))~c(ssb)))
    return(FLPar(a=max(exp(res[1])), b=-max(res[2]), rho=0))
	},
  # lower and upper limits for optim()
	lower=rep(-Inf, 3),
	upper=rep( Inf, 3)
	)

  ## model to be fitted
	model  <- rec~a*ssb*exp(-b*ssb)

	return(list(logl=logl, model=model, initial=initial))}

#### Fit
model(nsher)<-rickerAR1()
nsherAR1 <-fmle(nsher)

plot(nsherAR1)
```
The code is provided for demonstration purposes only as the used dataset is not adequate for all 3 parameters of the SR model to be estimated. 

Finally, a SR model where covariates (e.g. NAO index), can be used to model environmental effects on the stock recruitment relationship.
```{r, SR_custom_covars}
# Fit a custom SR model with covariates
nao     <-read.table(url("https://www.esrl.noaa.gov/psd/data/correlation/nao.data"),
  skip=1, nrow=62, na.strings="-99.90")
dnms    <-list(quant="nao", year=1948:2009, unit="unique", season=1:12, area="unique")
nao     <-FLQuant(unlist(nao[,-1]), dimnames=dnms, units="nao")

# include NAO as covar (note that it must be a FLQuants with a single component
# called “covar” that matches the year span of the data) and adjust the model.

nsherCovA <- nsher
nsherCovA <- transform(nsherCovA,ssb=ssb/1000,rec=rec/1000)

#### Modified so temperature affects larval survival
rickerCovA <- function(){
  logl <- function(a, b, c, rec, ssb, covar){
              loglAR1(log(rec), log(a*(1+c*covar[[1]])*ssb*exp(-b*ssb)))}

  initial <- structure(function(rec, ssb, covar) {
		# The function to provide initial values
    res  <-coefficients(lm(c(log(rec/ssb))~c(ssb)))
    return(FLPar(a=max(exp(res[1])), b=-max(res[2]), c=0.0))},

  # lower and upper limits for optim()
	lower=rep(-Inf, 3),
	upper=rep( Inf, 3))

	model  <- rec~a*(1+c*covar[[1]])*ssb*exp(-b*ssb)

	return(list(logl=logl, model=model, initial=initial))}


model(nsherCovA)<-rickerCovA()
covar(nsherCovA)<-FLQuants(covar=seasonMeans(trim(nao, year=dimnames(ssb(nsherCovA))$year)))
nsherCovA       <-fmle(nsherCovA,fixed=list(c=0))
plot(nsherCovA)
```


# References
Beverton, R.J.H. and Holt, S.J. (1957) On the dynamics of exploited fish populations. MAFF Fish. Invest., Ser: II 19, 533. ISBN: 1930665946

Needle, C.L. Recruitment models: diagnosis and prognosis. Reviews in Fish Biology and Fisheries 11: 95-111, 2002. DOI: https://doi.org/10.1023/A:1015208017674

Ricker, W.E. (1954) Stock and recruitment. J. Fish. Res. Bd Can. 11, 559-623. DOI: https://doi.org/10.1139/f54-039

Shepherd, J.G. (1982) A versatile new stock-recruitment relationship for fisheries and the construction of sustainable yield curves. J. Cons. Int. Explor. Mer 40, 67-75. DOI: https://doi.org/10.1093/icesjms/40.1.7


# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* ggplot2: `r packageVersion('ggplot2')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Nikolaos NIKOLIOUDAKIS**. Institute of Marine Research (IMR), Pelagic Fish Group, Nordnesgaten 33, P.O. Box 1870, 5817 Bergen, Norway. <http://www.imr.no/>
