---
title: "Running Medium Term Forecasts with **FLasher**"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
github_document:
  mathjax: TRUE
pdf_document:
  fig_width: 6 
  fig_height: 4 
tags: [FLR]
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE}
library(knitr)
source("R/ini.R")
```

# Introduction

This tutorial describes how Medium-Term Forecasts (MTF) can be performed using FLR. It uses the **FLasher** package for running projections, an updated version of the**FLash** package.

MTFs use the same engine as Short-Term Forecasts (STFs). However, there are some key differences between them.

* MTFs typically project over 5 to 10 years instead of the usual 3 years for a STF. Because of this increase in projection length it is necessary to include a stock-recruitment relationship to simulate the dynamics of the biological stock (an STF uses a constant recruitment assumption).
* MTFs may also have a more complicated projection control object because they can try to simulate management objectives (e.g. decreases in F over time).
* Finally, MTFs may also include consideration of uncertainty by including stochasticity in the projections.

Special attention must be paid to the conditioning and future assumptions of the stock.

This tutorial focuses on running projections with an *FLStock*. To see how **FLasher** can be used to simulate mixed fisheries with **FLBiol**s and **FLFisheries** see 
this [tutorial](http://www.flr-project.org/doc/Mixed_Fisheries_Projections_with_FLasher.html).

## Required packages

To follow this tutorial you should have installed the following packages:

- FLR: [FLCore](http://www.flr-project.org/FLCore/), [FLasher](http://www.flr-project.org/FLash/), [FLFishery](http://www.flr-project.org/FLFishery/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("FLCore"), repos="http://flr-project.org/R")
install.packages(c("FLasher"), repos="http://flr-project.org/R")
install.packages(c("FLFishery"), repos="http://flr-project.org/R")
```

```{r, pkgs}
# Load all necessary packages, trim pkg messages
library(FLCore)
library(FLasher)
```

# Introduction to Medium Term Forecasts

Running an MTF is similar to running an STF in that we need several components:

1. An *FLStock* object set up for the future (assumptions);
2. A stock-recrutiment relationship (SRR);
3. A projection control object;

However, there are some significant differences between an MTF and an STF:

i. An MTF is usually run for 5 to 10 years (an STF for 3 years);
ii. An MTF can use different target types (e.g. setting catch targets, not just F targets);
iii. A dynamic SRR should be used (the STF assumption of mean recruitment is not a good one for a projection of more than 3 years);
iv. We can include uncertainty in the initial *FLStock*, projected recruitment and target values.

In this tutorial we will build a 10 year projection, introduce a range of target types (including minimum and maximum target values, as well as relative target values), use a dynamic SRR and introduce uncertainty.

As ususal, we base the projections on plaice in the North Sea.

```{r, ple4}
data(ple4)
```

# Conditioning the projection

The first step is to condition the projection by making assumptions about the stock in the future, and to fit the SRR.

## Making the future stock

We use the _stf()_ function to set up our stock into the future. _stf()_ makes a lot of assumptions to set up a future stock. We may want to change some of these assumptions, but for the moment we will use the defaults.

```{r, stf_condition}
ple4_mtf <- stf(ple4, nyears = 10)
# Now the stock goes up to 2018
summary(ple4_mtf)
```
## The stock-recruitment relationship

In these examples we use a Beverton-Holt model (see the tutorial on fitting SRRs for more detail). The resulting SRR fit can be seen in `r fign("plotSRR")`.
Most of the projections will be performed by passing in a fitted *FLSR* object.
For more information on how to set the SR for projections with **FLasher** this [tutorial](http://www.flr-project.org/doc/Setting_Stock_Recruitment_in_FLasher_Projections.html).

```{r, fitSRR}
ple4_sr <- fmle(as.FLSR(ple4, model="bevholt"), control=list(trace=0))
```
```{r, plotSRR, fig.cap="Fitted Beverton-Holt stock-recruitment relationship for the *ple4* stock object"}
plot(ple4_sr)
```

# Example 1: Fbar targets

For a simple first example we will set the future F at F status quo and assume that F status quo is the mean of the last 4 years

```{r, ex1a}
f_status_quo <- mean(fbar(ple4)[,as.character(2005:2008)])
f_status_quo
```

Make the control `data.frame` including all the years of the projection (note that `FLash` used *quantity* and *val* as column names and `FLasher` uses *quant* and *value*)

```{r, ex1b}
ctrl_target <- data.frame(year = 2009:2018,
			                    quant = "f",
			                    value = f_status_quo)
```

Make the `fwdControl` object from the control `data.frame`

```{r, ex1c}
ctrl_f <- fwdControl(ctrl_target)
ctrl_f
```
An alternative method for making the *fwdControl* object is to pass in lists of target. This can often be simpler when making complicated control objects as we will see below.

```{r}
ctrl_f <- fwdControl(list(year=2009:2018, quant="f", value=f_status_quo))
ctrl_f
```

We have columns of *year*, *quant* (target type), *min*, *value* and *max* (and others not shown here yet). Here we are only using *year*, *quant* and *value*. We can now run _fwd()_ with our three ingredients. Note that the *control* argument used to be called *ctrl* in **FLash**. Also, with **FLasher** the *control* and *sr* arguments must be named.

```{r, ex1d}
ple4_f_sq <- fwd(ple4_mtf, control = ctrl_f, sr = ple4_sr)
# What just happened? We plot the stock from the year 2000.
plot(window(ple4_f_sq, start=2000))
```

The future Fs are as we set in the control object

```{r, ex1e}
fbar(ple4_f_sq)[,ac(2005:2018)]
```

What about recruitment? Remember we are now using a Beverton-Holt model.

```{r, ex1f}
rec(ple4_f_sq)[,ac(2005:2018)]
```

The recruitment is not constant but it is not changing very much. That's because the fitted model looks flat (`r fign("plotSRR")`).

# Example 2: A decreasing catch target {#ex2}

In this example we introduce two new things:

1. A new target type (catch)
2. A changing target value

Setting a catch target allows exploring the consequences of different TAC strategies. In this example, the TAC (the total catch of the stock) is reduced 10% each year for 10 years.

We create a vector of future catches based on the catch in 2008:

```{r, ex2a}
future_catch <- c(catch(ple4)[,"2008"]) * 0.9^(1:10)
future_catch
```

We create the *fwdControl* object, setting the target quantity to *catch* and passing in the vector of future catches

```{r, ex2b}
ctrl_catch <- fwdControl(list(year=2009:2018, quant = "catch", value=future_catch))
# The control object has the desired catch target values
ctrl_catch
```
We call `fwd()` with the stock, the control object and the SRR, and look at the results

```{r, ex2c}
ple4_catch <- fwd(ple4_mtf, control = ctrl_catch, sr = ple4_sr)
catch(ple4_catch)[,ac(2008:2018)]
plot(window(ple4_catch, start=2000))
```

The decreasing catch targets have been hit. Note that F has to be similarly reduced to hit the catch targets, resulting in a surge in SSB.

# Example 3: Setting biological targets

In the previous examples we have set target types based on the activity of the fleet (F and catch). We can also set biological target types. This is useful when there are biological reference points, e.g. Bpa.

Setting a biological target must be done with care because it may not be possible to hit the target. For example, even when F is set to 0, the stock may not be productive enough to increase its abundance sufficiently to hit the target.

There are currently three types of biological target available in `FLasher`: *SRP*, *SSB* and *biomass*. Of these, there are several flavours of *SSB* and *biomass* that differ in terms of timing.

The *SRP* target is the Stock Recruitment Potential *at the time of spawning*, i.e. if a stock spawns in the middle of the year, after the abundance has been reduced by fishing and natural mortality, this is the SRP at that point in time. At the moment, SRP is calculated as the mass of mature fish, i.e. SSB. If setting an *SRP* target, you must be aware of the timing of spawning and the timing of the fishing period.

Internally, **FLasher** attempts to hit the desired target in a time step by finding the appropriate value of F in that timestep. If the stock spawns before fishing starts, then changing the fishing activity in that timestep has no effect on the SRP at the time of spawning. It is not possible to hit the target by manipulating F in that timestep and **FLasher** gives up.

*SSB* is the Spawning Stock Biomass calculated as the total biomass of mature fish. The *biomass* is simply the total biomass of the stock. For the *SSB* and *biomass* targets, there are three different flavours based on timing:

* *ssb_end* and *biomass_end* - at the end of the time step after all mortality (natural and fishing) has ceased;
* *ssb_spawn* and *biomass_spawn* - at the time of spawning (mimics the `ssb()` method for *FLStock* objects);
* *ssb_flash* and *biomass_flash* - an attempt to mimic the behaviour of the original **FLash** package.

This last option needs some explanation. If fishing starts before spawning (i.e. the *harvest.spwn* slot of an *FLStock* is greater than 0) then the SSB or biomass at the time of spawning in that timestep is returned. If fishing starts after spawning, or there is no spawning in that time step (which may happen with a seasonal model), then the SSB or biomass at the time of spawning in the next timestep is returned.

However, this second case can be more complicated for several reasons. If there is no spawning in the next time step then we have a problem and *FLasher* gives up (F in the current timestep does not affect the SSB or biomass at the time of spawning in the current or next timestep). Additionally, if there is no next time step (i.e. we have reached the end of the projection) then *FLasher* gives up.

There is also a potential problem that if the fishing in the next timestep starts before spawning, the SSB or biomass at the time of spawning in the next timestep will be affected by the effort in the current timestep AND the next timestep. *FLasher* cannot handle this and weird results will occur (although it is an unusal situation).

For these reasons, it is better to only use the **FLash**-like target for annual models and when fishing and spawning happen at the same time in each year through the projection.

## Demonstrating the biological targets

Here we give simple demonstrations of the different types of biological targets using SSB. The results of using a biomass target will have the same behaviour. Only a 1 year projection is run.

The timing of spawning and fishing are controlled by the *m.spwn* and *harvest.spwn* slots. Our test *FLStock* object has *m.spwn* and *harvest.spwn* values of 0. This means that spawning and fishing happens at the start of the year and that spawning is assumed to happen before fishing.

### Targets at the end of the timestep

Here we set a target SSB for the end of the timestep

```{r, ex3a}
final_ssb <- 100000
ctrl_ssb <- fwdControl(list(year=2009, quant = "ssb_end", value=final_ssb))
ple4_ssb <- fwd(ple4_mtf, control=ctrl_ssb, sr = ple4_sr)
# Calculate the final SSB to check the target has been hit
survivors <- stock.n(ple4_ssb) * exp(-harvest(ple4_ssb) - m(ple4_ssb))
quantSums((survivors * stock.wt(ple4_ssb) * mat(ple4_ssb))[,ac(2009)])
```

### Targets at the time of spawning

If fishing occurs after spawning, the level of fishing will not affect the SSB or biomass at the time of spawning. This is currently the case because *m.spwn* and *harvest.spwn* have values of 0. The result is that the projection will fail with a warning (intentionally). We see this here.

```{r, ex3b, warning=TRUE}
spawn_ssb <- 100000
ctrl_ssb <- fwdControl(list(year=2009, quant = "ssb_spawn", value=spawn_ssb))
ple4_ssb <- fwd(ple4_mtf, control=ctrl_ssb, sr = ple4_sr)
# Using the `ssb()` method to get the SSB at the time of spawning, we can see that the projection failed
ssb(ple4_ssb)[,ac(2009)]
```

In the previous example, spawning happens at the start of the year. We can change this with the *m.spwn* slot. Natural mortality is assumed to happen continuously through the year. Therefore, if we set the *m.spwn* slot to 0.5, then half the natural mortality happens before spawning, i.e. spawning happens half way through the year. Similarly, the current value of *harvest.spwn* is 0, meaning that spawning happens before any fishing happens. If we set this value to 0.5 then half of the fishing mortality has occurred before spawning. With these changes, the example now runs.

```{r, ex3c}
m.spwn(ple4_mtf)[,ac(2009)] <- 0.5
harvest.spwn(ple4_mtf)[,ac(2009)] <- 0.5
spawn_ssb <- 100000
ctrl_ssb <- fwdControl(data.frame(year=2009, quant = "ssb_spawn", value=spawn_ssb))
ple4_ssb <- fwd(ple4_mtf, control=ctrl_ssb, sr = ple4_sr)
# We hit the target
ssb(ple4_ssb)[,ac(2009)]
```

At the moment `FLasher` calculates the SRP as SSB. This means that the *SRP* target type behaves in the same way as the *ssb_spawn* target.

```{r, ex3d}
srp <- 100000
ctrl_ssb <- fwdControl(data.frame(year=2009, quant = "srp", value=srp))
ple4_ssb <- fwd(ple4_mtf, control=ctrl_ssb, sr = ple4_sr)
# We hit the target
ssb(ple4_ssb)[,ac(2009)]
```

### **FLash**-like targets

As mentioned above, the `FLash`-like targets can have different behaviour depending on the timing of spawning and fishing. If fishing starts before spawning, the SSB or biomass at the time of spawning *in the current timestep* will be hit (if possible). This is demonstrated here.

```{r, ex3e}
# Force spawning to happen half way through the year
# and fishing to start at the beginning of the year
m.spwn(ple4_mtf)[,ac(2009)] <- 0.5
harvest.spwn(ple4_mtf)[,ac(2009)] <- 0.5
flash_ssb <- 150000
ctrl_ssb <- fwdControl(data.frame(year=2009, quant = "ssb_flash", value=flash_ssb))
ple4_ssb <- fwd(ple4_mtf, control=ctrl_ssb, sr = ple4_sr)
# Hit the target? Yes
ssb(ple4_ssb)[,ac(2009)]
```

However, if fishing starts after spawning, the SSB or biomass at the time of spawning *in the next timestep* will be hit (if possible). This is because fishing in the current timestep will have no impact on the SSB at the time of spawning in the current timestep.

```{r, ex3f}
# Force spawning to happen at the start of the year before fishing
m.spwn(ple4_mtf)[,ac(2009)] <- 0.0
harvest.spwn(ple4_mtf)[,ac(2009)] <- 0.0
flash_ssb <- 150000
ctrl_ssb <- fwdControl(data.frame(year=2009, quant = "ssb_flash", value=flash_ssb))
ple4_ssb <- fwd(ple4_mtf, control=ctrl_ssb, sr = ple4_sr)
# We did hit the SSB target, but not until 2010.
ssb(ple4_ssb)[,ac(2009:2010)]
```

## A longer SSB projection

Here we run a longer projection with a constant **FLash**-like SSB target. Spawning happens before fishing so the target will not be hit until the following year. 

```{r, ex3g}
# Force spawning to happen at the start of the year before fishing
m.spwn(ple4_mtf)[,ac(2009)] <- 0.0
harvest.spwn(ple4_mtf)[,ac(2009)] <- 0.0
future_ssb <- 200000
ctrl_ssb <- fwdControl(data.frame(year=2009:2018, quant = "ssb_flash", value=future_ssb))
ple4_ssb <- fwd(ple4_mtf, control = ctrl_ssb, sr = ple4_sr)
```

We get a warning about running out of room. This is because future stock object, *ple4_mtf*, goes up to 2018. When we set the SSB target for 2018, it tries to hit the final year target in 2019. The targets that were set for 2009 to 2017 have been hit in 2010 to 2018. However, we cannot hit the target that was set for 2018. This means that the returned value of F in 2018 needs to be discounted.

```{r, ex3h}
ssb(ple4_ssb)[,ac(2009:2018)]
fbar(ple4_ssb)[,ac(2009:2018)]
plot(window(ple4_ssb, start=2000, end=2017))
```

# Example 4:  Relative catch target

The examples above have dealt with *absolute* target values. We now introduce the idea of *relative* values. This allows us to set the target value *relative* to the value in another time step. 

We do this by using the *relYear* column in the control object (the year that the target is relative to). The *value* column now holds the relative value, not the absolute value.

Here we set catches in the projection years to be 90% of the catches in the previous year, i.e. we want the catche in 2009 to be 0.9 * value in 2008 etc.

```{r, ex4a}
ctrl_rel_catch <- fwdControl(
	data.frame(year = 2009:2018,
		   quant = "catch",
		   value = 0.9,
		   relYear = 2008:2017))
# The relative year appears in the control object summary
ctrl_rel_catch
```
We run the projection as normal

```{r, ex4b}
ple4_rel_catch <- fwd(ple4_mtf, control = ctrl_rel_catch, sr = ple4_sr)
catch(ple4_rel_catch)
catch(ple4_rel_catch)[,ac(2008:2018)] / catch(ple4_rel_catch)[,ac(2007:2017)]
```
```{r, ex4c, fig.cap="Relative catch example"}
plot(window(ple4_rel_catch, start = 2001, end = 2018))
```

This is equivalent to the catch example above (LINK) but without using absolute values.

# Example 5: Minimum and Maximum targets

In this Example we introduce two new things:

1. Multiple target types;
2. Targets with *bounds*.

Here we set an F target so that the future F = F0.1. However, we also don't want the catch to fall below a minimum level. We do this by setting a *minimum* value for the catch. 

First we set a value for F0.1 (you could use the `FLBRP` package to do this (LINK))

```{r, ex5a}
f01 <- 0.1
```

We'll set our minimum catch to be the mean catch of the last 3 years.

```{r, ex5b}
min_catch <- mean(catch(ple4_mtf)[,as.character(2006:2008)])
min_catch
```

To create the control object, we could make a `data.frame` with both target types and a *value* and a *min* column.
However, it is probably easier to use the *list* constructor when making the *fwdControl*.
Multiple lists can be passed in, each one representing a different catch. Here we pass in two lists, one for the F target and one for the minimum catch. Note the use of *min* in the second target list and that there is no *value*.

It is important that when running the projection, the bounding targets (the *min* and the *max*) are processed after the non-bounding targets. This should be sorted out by the *fwdControl* constructor.

```{r, ex5c}
ctrl_min_catch <- fwdControl(
    list(year=2009:2018, quant="f", value=f01),
    list(year=2009:2018, quant="catch", min=min_catch))
ctrl_min_catch
```

What did we create? We can see that the *min* column has now got some data (the *max* column is still empty) and the targets appear in the correct order. Now project forward

```{r, ex5e}
ple4_min_catch <- fwd(ple4_mtf, control = ctrl_min_catch, sr = ple4_sr)
fbar(ple4_min_catch)[,ac(2008:2018)]
catch(ple4_min_catch)[,ac(2008:2018)]
```

What happens? The catch constraint is hit in every year of the projection.
The projected F decreases but never hits the target F because the minimum catch constraint prevents it from dropping further.

```{r, ex5f, fig.cap="Example with a minimum catch bound and constant F target"}
plot(window(ple4_min_catch, start = 2001, end = 2018))
```

It is possible to also set a maximum constraint, for example, to prevent F from being too large.

# Example 6 - Relative targets and bounds

In this example we use a combination of *relative* targets and *bounds*.

This kind of approach can be used to model a recovery plan. For example, we want to decrease F to F0.1 by 2015 (absolute target value) but catches cannot change by more than 15% each year (relative bound). This requires careful setting up of the control object.

We make a vector of the desired F targets using the F0.1 we calculated above. We set up an F sequence that decreases from the current Fbar in 2008 to F01 in 2015, then F01 until 2018.

```{r, ex6a}
current_fbar <- c(fbar(ple4)[,"2008"])
f_target <- c(seq(from = current_fbar, to = f01, length = 8)[-1], rep(f01, 3))
f_target
```

We set maximum annual change in catch to be 10% (in either direction).

```{r, ex6b}
rel_catch_bound <- 0.10
```

We make the control object by passing two lists. The first list has the fixed F target, the second list has the relative minimum and maximum bounds on the catch (set using *relYear*, *min* and *max*).

```{r, ex6c}
ctrl_rel_min_max_catch <- fwdControl(
    list(year=2009:2018, quant="f", value=f_target),
    list(year=2009:2018, quant="catch", relYear=2008:2017, max=1+rel_catch_bound, min=1-rel_catch_bound))
ctrl_rel_min_max_catch
```

Run the projection:

```{r, ex6e}
recovery<-fwd(ple4_mtf, control=ctrl_rel_min_max_catch, sr=ple4_sr)
```

What happened? The F decreased and then remains constant, while the catch has changed by only a limited amount each year.

```{r, ex6f}
plot(window(recovery, start = 2001, end = 2018))
```

The minimum and maximum bounds on the catch are operational during the projection. They prevent the catch from increasing as well as decreasing too strongly, (supposedly) providing stability to the fishery.

```{r, ex6g}
catch(recovery)[,ac(2009:2018)] / catch(recovery)[,ac(2008:2017)]
```

# Projections with stochasticity 

So far we have looked at combinations of:

* Absolute target values;
* Relative target values;
* Bounds on targets, and
* Mixed target types.

But all of the projections have been deterministic, i.e. they all had only one iteration.
Now, we are going start looking at projecting with multiple iterations.
This is important because it can help us understand the impact of uncertainty (e.g. in the stock-recruitment relationship).

_fwd()_ is happy to work over iterations.
It treats each iteration separately. 
"All" you need to do is set the arguments correctly.

There are four main ways of introducing iterations into fwd():

1. Having existing iterations and variability in the *FLStock* that is to be projected.
2. By passing in residuals to the stock-recruitment function (as another argument to _fwd()_);
3. Through the control object (by setting target values as multiple values)
4. Including iterations in the parameters for the stock-relationship relationship.

You can actually use both of these methods at the same time.
As you can probably imagine, this can quickly become a bit complicated so we'll just do some simple examples to start with.

## Preparation for projecting with iterations

To perform a stochastic projection you need a stock object with multiple iterations.
If you are using the output of a stock assessment method, such as *a4a*, then you may have one already.
Here we use the *propagate()* method to expand the ple4 stock object to have 200 iterations.
Each iteration is the same but we could have imposed some variability across the iterations (e.g. across the mean weights at age).
We'll use the ten year projection as before (remember that we probably should change the assumptions that come with the _stf()_ method).

```{r, niter}
niters <- 200
ple4_mtf <- stf(ple4, nyears = 10)
ple4_mtf <- propagate(ple4_mtf, niters)
```

You can see that the 6th dimension, iterations, now has length 200:

```{r, prop}
summary(ple4_mtf)
```

## Example 7: Stochastic recruitment with residuals

There is an argument to _fwd()_ that we haven't used yet: *residuals*
   
This is used for specifying the recruitment residuals (*residuals*) which are multiplicative.
In this example we'll use the residuals so that the predicted recruitment values in the projection = deterministic recruitment predicted by the SRR model * residuals.
The residuals are passed in as an *FLQuant* with years and iterations.
Here we make an empty *FLQuant* that will be filled with residuals.

```{r, res}
rec_residuals <- FLQuant(NA, dimnames = list(year=2009:2018, iter=1:niters))
```
We're going to use residuals from the stock-recruitment relationship we fitted at the beginning.
We can access these using:

```{r, res2}
residuals(ple4_sr)
```

These residuals are on a log scale i.e.  log_residuals = log(observed_recruitment) - log(predicted_recruitment).
To use these log residuals multiplicatively we need to transform them with *exp()*:

We want to fill up our *multi_rec_residuals* *FLQuant* by randomly sampling from these log residuals.
We can do this with the *sample()* function.
We want to sample with replacement (i.e. if a residual is chosen, it gets put back in the pool and can be chosen again).

First we get generate the samples of the years (indices of the residuals we will pick).

```{r, res3}
sample_years <- sample(dimnames(residuals(ple4_sr))$year, niters * 10, replace = TRUE)
```

We fill up the **FLQuant** we made earlier with the residuals using the sampled years:

```{r, res4}
rec_residuals[] <- exp(residuals(ple4_sr)[,sample_years])
```

What have we got?

```{r, res5}
rec_residuals
```

It's an *FLQuant* of SRR residuals but what do those brackets mean?
The information in the brackets is the Median Absolute Deviation, a way of summarising the iterations. We have 200 iterations but don't want to see all of them - just a summary.

We now have the recruitment residuals.
We'll use the *ctrl_catch* control object we made earlier with decreasing catch. Note that this control object does not have iterations. The target values are recycled over the iterations in the projection.

We call *fwd()* as usual, only now we have a *residuals* argument.
This takes a little time depending on the number of iterations.

```{r, res6}
ple4_stoch_rec <- fwd(ple4_mtf, control = ctrl_catch, sr = ple4_sr, residuals = rec_residuals) 
```

What just happened? We can see that now we have uncertainty in the recruitment estimates, driven by the residuals.
This uncertainty feeds into the SSB and, to a lesser extent, the projected F and catch.

```{r, res7, fig.cap="Example projection with stochasticity in the recruitment residuals"}
plot(window(ple4_stoch_rec, start = 2001, end = 2018))
```

We can see that the projected stock metrics also have uncertainty in them.

```{r, res8}
rec(ple4_stoch_rec)[,ac(2008:2018)]
fbar(ple4_stoch_rec)[,ac(2008:2018)]
ssb(ple4_stoch_rec)[,ac(2008:2018)]
```

## Example 8: stochastic target values

In this example we introduce uncertainty by including uncertainty in our target values.
This example has catch as the target, except now catch will be stochastic.

We will use the *ctrl_catch* object from above.

```{r, stv1}
ctrl_catch
```

Let's take a look at what else is in the control object:

```{r, stv2}
slotNames(ctrl_catch)
```

The iterations of the target value are set in the *iters* slot.

```{r, stv3}
ctrl_catch@iters
```

What is this slot?

```{r, stv4}
class(ctrl_catch@iters)
dim(ctrl_catch@iters)
```

It's a 3D array with structure: target no x value x iteration.
It's in here that we set the stochastic projection values.
Each row of the *iters* slot corresponds to a row in the control *data.frame* (the *target* slot).

One option for adding iterations to the control object is to manually adjust the *iters* slot to have the required number of iterations and fill it up accordingly.

However, it is easier to use the *list* constructor for the control object.
Here we generate random values, sampled from a lognormal distribution, for the catch based on the *future_catch* object we created earlier.
This is a vector of length 2000. Passing this into the constructor with 10 years makes a control object with 200 iterations.

```{r}
stoch_catch  <- rlnorm(10*niters, meanlog=log(future_catch), sdlog=0.3)
length(stoch_catch)
ctrl_catch_iters <- fwdControl(list(year=2009:2018, quant="catch", value=stoch_catch))
```

The control object has 200 iterations. The variance in the values can be seen.

```{r}
ctrl_catch_iters
```

We project as normal using the deterministic SRR.

```{r, stv10}
ple4_catch_iters <- fwd(ple4_mtf, control=ctrl_catch_iters, sr = ple4_sr)
```

What happened?

```{r, stv11}
plot(window(ple4_catch_iters, start = 2001, end = 2018))
```

The projected catches reflect the uncertainty in the target.

```{r, stv12}
catch(ple4_catch_iters)[,ac(2008:2018)]
```

## Example 9: Including iterations in the stock-recruitment relationship

In this example we include stochasticity in the SR relationship.
The SR model fitted above is deterministic in that the parameters only have 1 iteration.

```{r}
params(ple4_sr)
```

The parameters are specified as an *FLPar*. To include stochasticity in the SR parameters it is necessary to make a similar *FLPar* which has iterations.


```{r}
sr_iters <- FLPar(NA, dimnames=list(params=c("a","b"), iter=1:niters))
```

We then need to generate some stochastic values of the _a_ and _b_ parameters.
These could be sampled from a distribution based on the SR model fit (see an example in this [tutorial](http://www.flr-project.org/doc/Setting_Stock_Recruitment_in_FLasher_Projections.html)).
Here we just pull some numbers from a distribution without worrying about covariance.
The danger of doing is that we can end up with pairs of values for _a_ and _b_ which are stupid.

```{r}
aiters <- rlnorm(niters, meanlog=log(params(ple4_sr)["a"]), sdlog=0.5)
biters <- rlnorm(niters, meanlog=log(params(ple4_sr)["b"]), sdlog=0.01)
```

We then put this into the *FLPar*.

```{r}
sr_iters["a"] <- aiters
sr_iters["b"] <- biters
sr_iters
```
The parameters are now all set up.
We could put these into the *FLSR* object. Instead we use a different method for specifying the SR relationship, using the model name and the parameters.
This projection only has stochasticity in the SR params.

```{r}
ple4_sr_iters <- fwd(ple4_mtf, control=ctrl_catch, sr = list(model="bevholt", params=sr_iters))
plot(window(ple4_sr_iters, start = 2001, end = 2018))
```

## Example 10: A projection with stochastic targets and recruitment

In this example we use the stochastic catch target, residuals and stochasticity in the SR parameters:

```{r, stv14}
ple4_iters <- fwd(ple4_mtf, control=ctrl_catch_iters, sr = list(model="bevholt", params=sr_iters), residuals = rec_residuals)
```

What happened?

```{r, stv15}
plot(window(ple4_iters, start = 2001, end = 2018))
```

We have a projection with stochastic target catches and recruitment. 

```{r, stv16}
catch(ple4_catch_iters)[,ac(2008:2018)]
rec(ple4_catch_iters)[,ac(2008:2018)]
```

Super.

# TO DO

## Alternative syntax for controlling the projection

SOMETHING ON CALLING FWD() AND SPECIFYING TARGETS AS ARGUMENTS

## Notes on conditioning projections

SOMETHING ON FWD WINDOW

# References


# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLasher: `r packageVersion('FLasher')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Finlay Scott**. European Commission, DG Joint Research Centre, Directorate D - Sustainable Resources, Unit D.02 Water and Marine Resources, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>
