---
title: Virtual Population analysis using eXtended Survivor Analysis
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document
tags:
license: Creative Commons CC-BY SA
bibliography: bibliography.bib
---

```{r, ini, echo=FALSE, results='hide', message=FALSE, warnings=FALSE, cache=FALSE}
library(knitr)
source("R/ini.R")
```

## Required packages

To follow this tutorial you should have installed the following packages:

* FLR: [FLCore](http://www.flr-project.org/FLCore/), [FLAssess](http://www.flr-project.org/FLAssess/),
       [FLXSA](http://www.flr-project.org/FLXSA/), [ggplotFL](http://www.flr-project.org/ggplotFL/)
* CRAN: [reshape]
You can do so as follows,

```{r echo=TRUE, eval=FALSE}
install.packages(c("FLCore", "FLAssess", "FLXSA"), repos="http://flr-project.org/R")
```

```{r echo=TRUE, eval=TRUE}
# This chunk loads all necessary packages, trims pkg messages
library(FLCore)
library(FLAssess)
library(FLXSA)
```

```{r echo=FALSE, eval=TRUE, message = FALSE}
library(plyr)
library(ggplot2)

setMethod('plot', signature(x='FLXSA', y='missing'),function(x){
  
  ggplot(subset(x@diagnostics,age>0))+
    geom_point(aes(as.numeric(yrcls),nhat,size=0),shape=1,alpha=0)+
    geom_point(aes(as.numeric(yrcls),nhat,fill=factor(age),size=w),shape=21,
               data=subset(x@diagnostics,age>0&w>0))+
    geom_vline(aes(xintercept=range(x)["maxyear"]-x@control@shk.yrs-0-5))+
    scale_x_discrete(name="") +
    facet_grid(source~age)+
    theme_bw()+theme(legend.position="none")})
```

# Introduction

## What is VPA
  Virtual  population  analysis  (VPA)  is  a  modeling  technique  commonly  used  in  fisheries  science  for
  reconstructing  the  historical  population  structure  of  an age structured fish  stock  using 
  information  on  the  deaths  of individuals in each time step. 
  The time steps are typically, though not necessarily, annual and the deaths are
  usually partitioned into mortality due to fishing and natural mortality. In some instances natural mortality
  may  be  further  partitioned  into  predation  mortality  and  mortality  from  other  causes,  such  as  disease,
  senesence etc.

  VPA is the most commonly used term to refer to cohort reconstruction techniques used in fisheries.  It is
  virtual  in the sense that  the population size is not observed or measured directly but is inferred or backcalculated
  to have been a certain size in the past.  Several  different software implementations of cohort
  reconstruction for fish populations exist including ADAPT which is often used in Canada and the USA and
  XSA [@shepherd1999] which is commonly used in Europe. The back-calculations in these implementations work
  the same way but they differ in the statistical methods used for "tuning" to indices of population size.
  Tuning  refers  to  the  use  of  auxilliary  information  to  determine  the  terminal  fishing  mortalities  and
  population numbers. Most tuning approaches involve a regression of fishing mortality against fishing effort
  to estimate  population abundance  at  age through an iterative  convergence  to some  threshold criterion.
  Relatively simple techniques, the Laurec-Shepherd method [@pope1985] for example, have been shown to work
  well with simulated data but there is little theoretical work to justify or validate these approaches [@quinn1999].

  A number of assessment methods are made available in FLR as well as the basic VPA tools to enable you to
  develop your own assessment methods. In this tutorial we will cover the basic VPA tools, simple methods
  for tuning a VPA and finally show how to run **FLXSA**. 

## Stock assessment methods within the FLR package structure
  The package **FLAssess** contains the basic class for age and biomass based stock assessments. 
  It provides a standard class, **FLAssess**, for data input,  
  stock status estimation and diagnostic inspection. The **FLAssess** package has a
  variety of uses. It can be applied within a stock assessment working group setting or, alternatively, as part of
  the management procedure in a formal Management Strategy Evaluation (MSE).
  FLAssess provides a common interface for existing stock assessment methods (e.g. XSA)
  allowing methods to be used interchangeably. It also includes various methods of general use such as setting 
  up a short-term forecast (`stf`), running VPAs (`VPA` or `SepVPA`) and calculating F
  from  catches.  There  are  several  steps  to  be  completed  when  conducting  an  assessment.  This  tutorial 
  considers only the process of running `VPA` and `FLXSA` stock assessment model.  
  Additional  tutorials are available, 
  e.g. [A quick introduction to FLR](http://doc.flr-project.org/A_quick_introduction_to_FLR.html), 
     [An overview of the FLCore classes](http://doc.flr-project.org/An_overview_of_the_FLCore_classes.html),
     [Loading your data into FLR](http://doc.flr-project.org/Loading_your_data_into_FLR.html).
   

  We will start  by importing the data sets
  for the North Sea Plaice stock and the fishery independent abundance indices. 
  We will use these example data sets for all of the examples in this tutorial. 

```{r echo=TRUE, eval=TRUE}
data(ple4)
data(ple4.indices)
```

  The North Sea Plaice  `FLStock` object  already has values estimated for harvest and stock numbers.  We
  should remove these first and replace them with `NA`.

```{r echo=TRUE, eval=FALSE}  
harvest(ple4)[] <- NA
stock.n(ple4)[] <- NA
```

 
 We should note at this point that the example below should not be considered the definitive assessment for
 the North Sea Plaice. We provide this example merely to show the procedure for conducting assessments using
FLR.

# The VPA method
  The  `VPA` method implements Pope's Virtual  Population Analysis (VPA).  It  is called with the command
  VPA which returns an object of class `FLVPA` that is itself an extension of the `FLAssess` class. 
  The  `VPA` method estimates population numbers and fishing mortalities at  age by back-calculating values
  down each cohort. To do this, the method requires initial values of harvest for the terminal age and terminal
  year in the `FLStock` object. These terminal values must be specified by the user prior to running the `VPA`.
  The arguments to the `VPA` method are the `FLStock` object for which values are to be calculated and two
  optional arguments.

  The range method will show details of the age and year range of the `ple4`  `FLStock` object. We can use
  this information to manually specify the terminal values in the harvest slot. In this instance we will set these  
  values to 1.0. Remember to convert the values to be of type character when indexing the `FLQuants`.

```{r echo=TRUE, eval=TRUE}   
harvest(ple4)[ac(range(ple4)["max"]), ]     <- 1
harvest(ple4)[, ac(range(ple4)["maxyear"])] <- 1

ple4.vpa <- VPA(ple4, fratio = 1, fit.plusgroup = T)
ple4.new <- ple4 + ple4.vpa

## Have a look in stock number ##
stock.n(ple4.vpa)[, ac(2005:range(ple4)["maxyear"])]

## Have a look in fishing mortality ##
harvest(ple4.vpa)[, ac(2004:range(ple4)["maxyear"])]

## Plot results ##
plot(FLStocks(ple4=ple4, vpa=ple4.new))
plot(FLStocks(vpa=ple4.new))
```

  The estimated population numbers and fishing mortality values at age from the `VPA` are now available in
  the returned object. Note that the terminal values for fishing mortalty are the user defined values that were
  specified prior to running the `VPA`.

# A simple method for tuning a VPA
  As  noted  above  the  VPA method  requires  user  defined  terminal  estimates  of  fishing  mortality.  This
  dependency limits the usefulness of the method since it is often the most recent, terminal, estimates that are
  of most concern to fishery managers. Additional catch at age and effort information, derived either from a
  sub component of the fishery or from a fishery independent source such as a research survey,  can be used to
  'tune'  the assessment,  as described above,  and thereby obtain better estimates of fishing mortality and
  stock numbers in the most recent years. Several so-called ad hoc techniques for tuning a VPA have been 
  developed. A relatively simple technique that has been widely used is the Laurec Shepherd method. This
  method can be easily implemented in FLR using the basic VPA tools that  are provided in the  **FLAssess**
  package. 
  
  The example shown below is a simple implementation that allows for a single tuning fleet. With a little extra
  effort it could be easily extended to accomodate multiple tuning fleets. The technical details of the method
  are not explained here. 
 
```{r echo=TRUE, eval=TRUE}
# Define Laurec-Sheperd function #

lsm <- function(stock, index, fratio = 1, fit.plusgroup = T) {
  harvest(stock)[, ac(range(stock)["maxyear"])] <- 0.5
  diff <- 1
  while (diff > 1e-06) {
    stock <- stock + VPA(stock, fratio = fratio)
    ages  <- range(index)["min"]:range(index)["max"]
    yrs   <- range(index)["minyear"]:range(index)["maxyear"]
    stk   <- trim(stock, year = yrs, age = ages)
    Cp    <- catch.n(index)/catch.n(stk)
    q     <- sweep(Cp * harvest(stk), 2, effort(index), "/")
    gmq   <- apply(q, 1, function(x) exp(mean(log(x), na.rm = T)))
    mFp   <- gmq * c(apply(effort(index), 1, mean))
    Fr    <- mFp * (apply(Cp, 1, mean, na.rm = T))^-1
    Fnew  <- c(Fr, rep(Fr[ac(max(ages)), ], 2))
    diff  <- sum(abs(harvest(stock)[, ac(range(stock)["maxyear"])] -
                        Fnew))
    harvest(stock)[, ac(range(stock)["maxyear"])] <- c(Fnew)
    }
  res <- VPA(stock, fratio = fratio, fit.plusgroup = fit.plusgroup)
  index.res(res) <- FLQuants(q)
  return(res)
  }
```

 
  The new Laurec-Shepherd function can now be called without having to specify terminal  values in the
  harvest  slot.  The arguments  to the  `VPA` method  are   also formally declared as  arguments  to our new 
  function. Note that the function returns an object of class  `FLVPA` that has been created from a call to the
  `VPA` method and that the catchability residuals are stored in the `index.res` slot of the returned object. 


```{r echo=TRUE, eval=TRUE}

harvest(ple4)[] <- NA
stock.n(ple4)[] <- NA

ple4.LSvpa <- lsm(ple4, ple4.indices[[1]], fratio = 1, fit.plusgroup = T)

ple4.new2 <- ple4 + ple4.LSvpa

stock.n(ple4.LSvpa)[, ac(2005:range(ple4)["maxyear"])]
harvest(ple4.LSvpa)[, ac(2004:range(ple4)["maxyear"])]

# Compare the results with previous fits.
plot(FLStocks(vpa=ple4.new2))
```

# `FLXSA`
  The Laurec-Shepherd method above is a relatively simple technique for tuning a VPA. XSA is a  
  more sophisticated  method that uses information on individual cohort sizes to estimate survivors at each age in
  the terminal population. Although the modelling approach is more involved the method requires the same 
  input of catch numbers at age and indices of catch per unit effort and it retains at its core the basic `VPA`
  method. The details of the XSA method are too complex to show here, or to code individually as we have 
  for the Laurec-Shepherd approach. Instead  the  `FLXSA` method has been developed as an additional package
  to **FLAssess**.

## The `FLXSA control` object
  The `FLXSA.control` object contains all of the user defined model settings for running an XSA analysis.
  It can be created in several different ways. The simplest method is to accept all of the default settings by
  calling the FLXSA.control function without any extra arguments: 

```{r echo=TRUE, eval=TRUE}
  FLXSA.control()
```

 Alternatively the default  settings can be over-written by specifying values at  the point of creation or by
overwriting them afterwards.

```{r echo=TRUE, eval=TRUE}
  ctrl <- FLXSA.control(maxit = 50, qage = 8)
  ctrl <- FLXSA.control()
  slot(ctrl, 'qage')  <- as.integer(8)
  slot(ctrl, 'maxit') <- as.integer(50)
```

Note that in the example above, when modifying the control object after creation, it is necessary to coerce
the values 8 and 50 to type integer. This is because the default type numeric cannot be used in this slot. Such 
coercion  is  not  necessary  when  using  the  `FLXSA.control` function  as  this  check  is  performed
internally by the function. You can use the `getSlots` function to determine the class of object associated
with any given slot.

```{r echo=TRUE, eval=TRUE}
 xsa.control <- FLXSA.control(maxit = 50, fse = 2.5)
 ple4.xsa    <- FLXSA(ple4, ple4.indices, xsa.control)
 ple4.xsa.t1 <- FLXSA(ple4, ple4.indices[[1]], xsa.control)
```
 
Once the control object has been created, the XSA analysis can be run as a one-line command. The
```FLXSA` function returns an object of class `FLXSA` which extends the `FLAssess` class. 
The `FLXSA` object 
contains all  of the information in the  `FLAssess` class plus additional  information specific to the XSA
assessment method, such as the survivors estimates  and their  internal  and external  standard errors.  The 
control object used for the assessment is also stored in the returned `FLXSA` object to provide a record of
what settings were used for that particular run. All of the settings in the returned control object will remain 
the same except for the `maxit` slot that contains the maximum number of iterations for the analysis. This
value will  be overwritten with the actual  number of iterations taken to reach convergence,  if indeed the 
model had converged before the maximum number initially specified.



## XSA Results
Appart from the model diagnostics, the `FLXSA` method returns two important results, namely the estimated
values of fishing mortality and population numbers at age. These are returned as `FLQuants` and are stored
in the `harvest` and `stock.n` slots, respectively,  of the `FLXSA` object. These estimated values can be
very easilly read back into an `FLStock` object using the `+` operator. Once the results have been read back
into a  FLStock object we can look at some of the key information such as SSB, recruitment and mean
fishing mortality values. But before concentrating too much on the results of the assessment it is advisable to 
first investigate some of the model diagnostics.

```{r echo=TRUE, eval=TRUE}
ple4.new  <- ple4 + ple4.xsa
ple4.ssb  <- ssb(ple4.new)
ple4.rec  <- rec(ple4.new)
ple4.fbar <- fbar(ple4.new)
```

## XSA Diagnostics
There are many diagnostic checks that one might be interested in conducting to examine the model fit. The
first might be to see if the model has reached convergence within the specified number of iterations. 

```{r echo=TRUE, eval=TRUE}
 slot(slot(ple4.xsa, "control"), "maxit")
```

Additionally one  can check for  discrepancies  between the internal  and external  standard  errors  of the
survivors estimates. Very often plots of the catchability residuals are made to inspect for any obvious trends
or departures from the assumption of constant catchability over time.  Some examples of these plots and
details of their creation from FLR objects are provided below but you should also consult the tutorial  on
lattice plotting and advanced graphics for FLR to see examples of other ways to graphically display your
data.



There are several ways to access diagnostic information about your fitted XSA model. The easiest is perhaps
to use the  `diagnostics` function, which will  replicate the diagnostic output produced by the original 
VPA suite (developed in the early 1990's). Note that this function merely outputs the results to the screen
and no object  is created by the method. The function was created to allow the user to cut and paste the
information from the console to a report.  The output  can be quite  long,  particularly if  the assessment
comprises a large number of ages and many tuning indices. The standard output can be divided roughly into
eight sections each providing different information about the model and the fit. These sections comprise the
model dimensions; parameter settings; regression weights; the estimated fishing mortalities and population
numbers for the last 10 years; the aggregated survivors estimates; the log catchability residuals for each of
the tuning indices  and finally the individual  survivors  estimates  for  each year-class  represented in the
terminal year. 

In order to make this document more readable we will print out only a few sections of the diagnostic output
at a time. We can do this by passing a vector of `TRUE` and `FALSE` values to the sections argument of the `diagnostics` method. By default all sections are set to `TRUE` so that all of the information is output to the screen. In order to reduce the quantity of output further  
we will  run a new XSA for a reduced number of ages and with only one tuning index and will start by
outputting only the dimension information and the parameter settings from our diagnostics.

```{r echo=TRUE, eval=TRUE}
 ple4.xsa2 <- FLXSA(trim(ple4, age = 1:4), ple4.indices[[3]],
     xsa.control)
 diagnostics(ple4.xsa2, sections = c(T, T, rep(F, 6)))
```

Next we can output the regression weights and the fishing mortalities and population numbers for the last 10
years and also the aggregated survivors estimates.

```{r echo=TRUE, eval=TRUE}
 diagnostics(ple4.xsa2, sections = c(F, F, T, T, T, T, F, F))
```


And finally we can output the catchability residuals and the individual survivors estimates. Note that very
little thought went into the parameter settings for this particular model  fit  so please don't  interrogate the
output presented here too closely. Also note that we do not normally expect  the diagnostics output to be
broken up as we have here. We present it in this way purely to make it more presentable in this document. 
By default all sections are set to  `TRUE ` so it is very likely that you won't need to give this argument at all 
when calling the diagnostics method.

```{r echo=TRUE, eval=TRUE}
 diagnostics(ple4.xsa2, sections = c(F, F, F, F, F, F, T, T))
```


Remember that the diagnostics method will only output text to the console, enabling you to copy and paste
the output to a report or other document. If you want to access the diagnostic data you will need to access
the specific slots of the returned `FLXSA` object. The information that you will requie is contained in various 
slots. The individual estimates of population number from each source (ie. tuning series  and F shrinkage)
and their individual weightings are stored as a dataframe in the `diagnostics` slot of the returned object.
Other  slots  contain the internal  and external  standard  errors;  the log catchability residuals.  For a more
thorough description of the XSA diagnostics you should consult the VPA users manaual.

## Plotting Diagnostics 
Very often the quickest and simplest way to determine the fit of the model is through visual inspection of the
various diagnostic outputs.

The default plot for '''FLXSA''' class shows the weight given to each of the indices, incluiding the shrinkage, 
to estimate total numbers at age along ages and years. The size of the bubbles in the plot is proportional to the 
weight given to the index to estimate the terminal numbers at age. The rows corresponds with the indices used and the 
columns with age classes.  The y axis represent the estimate of numbers at age obtained from each index.

```{r echo=TRUE, eval=TRUE}
 plot(ple4.xsa2)
```


Below we provide examples of how to extract the relevant information from the
return `FLXSA` object and to plot it using a variety of lattice functions available to `R`. 
We start by plotting the
log catchability residuals at age from each of the three tuning series. The data are stored as an `FLQuants` 
object  in the index.res slot of the `FLXSA` object.  First we need to assign names to each of the `FLQuant` 
objects so we know which fleet they represent.

```{r echo=TRUE, eval=TRUE}
 names(ple4.xsa@index.res) <- names(ple4.indices)
 pfun <- function(x, y, ...) {
     panel.xyplot(x, y, ...)
     panel.loess(x, y, ...)
     panel.abline(h = 0, col = "grey", lty = 2)
 }
 plot(xyplot(data ~ year | ac(age) + qname, data = index.res(ple4.xsa),
     panel = pfun))
```

A simple comparison of the terminal year survivors estimates can be obtained from the information stored in
the diagnostics slot of the `FLXSA` object. In the following example we first extract the information relevant
to the survivors estimates in the final year and store it as a temporary object. The weights values contained
in this data set are the raw fleet based weights that have been calculated from the standard errors of the fleet 
based survivors estimates at each age in the cohort. To aid visualisation and to see the relative contribution 
of each fleets estimate to the final estimated value of survivors we re-scale the weights to a maximum value
of 1 and plot both the fleet based survivors estimates from each fleet and their scaled weight. The results
show relatively consistent estimates of survivors from all fleets across most ages. The scaled weights show
the some series to have the greatest influence on the terminal estimates at the younger ages whilst others 
have greater influence at older ages and that throughout all ages F shrinkage recieves very little weighting.

```{r echo=TRUE, eval=TRUE}
 diag <- slot(ple4.xsa, "diagnostics")[is.element(slot(ple4.xsa,
     "diagnostics")$year, 2008), ]
 diag <- cbind(diag, w.scaled = diag$w/rep(tapply(diag$w, diag$yrcls, sum),
     c(table(diag$yrcls))))
 nplot <- barchart(ac(yrcls) ~ nhat, groups = source, data = diag,
     col = grey(c(0.1, 0.6, 0.3, 0.8)), main = "N Estimates",
     ylab = "Year Class", key = list(x = 0.3, y = 0.25, text = list(legend =
rev(c("BTS-Isis",
         "BTS-Tridens", "fshk", "SNS"))), rectangles = list(col =
grey(rev(c(0.1,
         0.6, 0.3, 0.8))))))
 wplot <- barchart(ac(yrcls) ~ w.scaled, groups = source, data = diag,
     col = grey(c(0.1, 0.6, 0.3, 0.8)), main = "Scaled Weights",
     ylab = "", xlab = "Relative Weight")
 print(nplot, position = c(0, 0, 0.5, 1), more = TRUE)
 print(wplot, position = c(0.5, 0, 1, 1))
```

## Sensitivity to different model settings.
The simplified calling format of  `FLXSA` makes it very easy to run multiple analyses to investigate model
sensitivity to parameter settings. A wide variety of such investigations are possible. In this simple example 
we will look at the effect that different F shrinkage standard errors have on the terminal estimates of fishing
mortality. We start by creating a vector of F shrinkage values to be used in the anlyses and by creating an
FLQuant with sufficient dimensions to store the results. To do this we use the propagate function to 
extend an FLQuant in the 6th  dimension by the number of runs that we are going to perform. The estimates
of fishing mortality for each XSA run are stored in the  FLQuant using the 6th
 dimension to hold each iteration. The results show little sensitivity to 
 increasing F shrinkage values at values between 1.0 and 2.5 .  

```{r echo=TRUE, eval=TRUE} 
 fsevals <- seq(0.5, 2.5, by = 0.5)
 res <- propagate(harvest(ple4), length(fsevals))
 for (i in 1:length(fsevals)) {
     xsa.control <- FLXSA.control(fse = fsevals[i])
     iter(res, i) <- harvest(FLXSA(ple4, ple4.indices, xsa.control))
 }
 plot(xyplot(data ~ year | age, groups = iter, data = res, type = "l",
     col = "black", xlim = c(1990:2010)))
```

## Retrospective Analyses
An important diagnostic check is to see how the estimated values vary as the time series of the input data
changes.  We can make use of existing `R` functions to apply the same assessment model  to successively
truncated the time series of input data.  In this example we are using  `window` to truncate the  `FLStock` 
object to the specified year range, the `+` operator to pass the results of the XSA into the `FLStock` object
and the  `tapply` function to perform this action over the year range `2004:2008`. Note that  the resulting
object, called `ple4.ret`, is of class `FLStocks` ie. a list of `FLStock` objects, each one having a separate
year range.

```{r echo=TRUE, eval=TRUE}
retro.years <- 2004:2008
ple4.retro <- tapply(retro.years, 1:length(retro.years), function(x){
	window(ple4,end=x)+FLXSA(window(ple4,end=x),ple4.indices)
})

# coerce into FLStocks object
ple4.retro <- FLStocks(ple4.retro)
# full retrospective summary plot
ple4.retro@names=ac(c(retro.years))###Add years to legend
plot(ple4.retro)
```


# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* FLXSA: `r packageVersion('FLXSA')`
* FLAssess: `r packageVersion('FLAssess')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Dorleta Garcia** AZTI. Marine Reserach Unit. Txatxarramendi Ugartea z/g, 48395, Sukarrieta, Basque Country, Spain.

**Alessandro MANNINI**. European Commission, DG Joint Research Centre, Directorate D - Sustainable Resources, Unit D.02 Water and Marine Resources, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>


# References
