---
title: PLOTTING FLR OBJECTS USING LATTICE
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document
tags:
license: Creative Commons CC-BY SA
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE}
# This chunk set the document environment, so it is hidden
library(knitr)
opts_chunk$set(fig.align='center',
  message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE)
options(width=50)
set.seed(1423)
```

The lattice package (Sarkar, 2008) improves on base R graphics by providing better defaults and the ability to easily display multivariate relationships. In particular, the package supports the creation of trellis graphs - graphs that display a variable or the relationship between variables, conditioned on one or more other variables. Lattice methods (e.g. xyplot(), bwplot(), dotplot()) are available for FLR classes in FLCore, and the standard plots (plot()) in FLR are lattice-based.

## Required packages

To follow this tutorial you should have installed the following packages:

- CRAN: [lattice](https://cran.r-project.org/web/packages/lattice/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/)

You can do so as follows:
```{r, eval=FALSE}
install.packages(c("lattice"))
install.packages(c("FLCore"), repos="http://flr-project.org/R")
```

Load necessary packages:
```{r, pkgs}
library(FLCore)
library(lattice)
```

# GRAPH TYPES AVAILABLE

The typical formula for lattice plots is:

**graph_type(formula, data=FLQuant)**

where *graph_type* is selected from the list below. Formula specifies the variable(s) to display and any conditioning variables. For example  ~data|A means display numeric variable for each level of factor A. data~A means display the relationship between numeric variables for factor A levels. ~x means display numeric variable x alone. 

The following lattice graph options could be used for FLR objects: 

Graph Type | Description        | Formula
-----------|------------------- |---------
bwplot     | boxplot            | data~A
dotplot    | dotplot            | data~A
histogram  | histogram          | ~data|A
xyplot     | scatterplot        | data~A
wireframe  | 3D wireframe graph | data~A*B
bubbles    | bubble plot        | A~B

In case of lattice based *plot* simply the name of the object could be used, e.g. *plot(ple4)*

# STANDARD LATTICE BASED FLR PLOTS

The standard plot() function in FLR returns lattice-based plots for FLR objects.

```{r fig1}
## Import your data from FLR
#place in the North Sea
data(ple4)
data(nsher)

# And the standard plots in FLR are lattice-based
# plot for FLQuant
plot(catch.n(ple4))
```


```{r fig2}
# Plot FLStock
plot(ple4)
```

```{r fig3}
# Plot FLSR
plot(nsher)
```

# PLOTTING FLR OBJECTS USING LATTICE


```{r fig4}
# The conversion to data.frame requires columns names for dims and 'data'. 
#The following names of FL object are used
head(as.data.frame(catch(ple4)))
```


## XYPLOTS for FLR

```{r fig5}
xyplot(data~year, data=catch(ple4), type='b', pch=19)

```


```{r fig6}
# values in FLQuant turn into a 'data' column to use in formulas
xyplot(data~year, groups=age, data=catch.n(ple4), type='l', ylab='Catch (t)')

```

```{r fig7}
# values in FLQuant turn into a 'data' column to use in formulas
xyplot(data~year|age, data=catch.n(ple4), type='l', ylab='Catch (t)')

```

```{r fig8}
# Methods are also available for FLQuants, with elements names' variable 'qname'
xyplot(data~year|qname, data=FLQuants(SSB=ssb(ple4), Yield=catch(ple4)))
```

```{r fig9}
# Methods are also available for FLQuants, with elements names' variable 'qname'
xyplot(data~year|qname, data=FLQuants(SSB=ssb(ple4), Yield=catch(ple4)))
```


## BWPLOTS

```{r fig10}
# bwplot
bwplot(data~year, rlnorm(200, fbar(ple4), 0.15),scales=list(x=list(rot=90)))

```

## DOTPLOTS

```{r fig11}

dotplot(data~year, groups=age, harvest(ple4),
        scales=list(x=list(rot=90)),
        	ylab="F at age")

```

## HISTOGRAM

```{r fig12}
histogram(~data|year, catch.n(ple4))

```




## WIREFRAME

```{r fig14}
wireframe(data~age*year, data=harvest(ple4),zlab="F")
```


# FLR specific plots BUBBLES

The bubble plots are created specially for FLR and are used to analise data by age classes. 

```{r fig15}
bubbles(age~year, data=catch.n(ple4), bub.scale=5)

```
One could also specify parts of the object of interest, e.g. select age classes between 5 and 10 as following:

```{r fig16}
bubbles(age~year, data=catch.n(ple4)[5:10,], bub.scale=10)

```

# References

[Quick Lattice Graphs](http://statmethods.net/advgraphs/trellis.html)  
[A quick introduction to FLR](http://doc.flr-project.org/A_quick_introduction_to_FLR.html)

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* lattice: `r packageVersion('lattice')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission Joint Research Centre (JRC), Institute for the Protection and Security of the Citizen (IPSC), Maritime Affairs Unit, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>
