---
title: TITLE
author: NAME (INSTITUTION) <email@place.country>
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Vignette Title}
  \usepackage[utf8]{inputenc}
tags:
abstract:
license: Creative Commons CC-BY SA
---
 

```{r, ini, echo=FALSE, results='hide', message=FALSE}
library(rmarkdown)
options(width=50)
set.seed(1423)
```

```{r, pkgs, echo=FALSE, message=FALSE}
library(FLCore)
```


```{r}
```



# References

https://creativecommons.org/licenses/by-sa/4.0

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* **Compiled**: `r date()`
* **Git Hash**: `r system("git log --pretty=format:'%h' -n 1", intern=TRUE)`

## Author information

**Iago MOSQUEIRA**. European Commission Joint Research Centre (JRC), Institute for the Protection and Security of the Citizen (IPSC), Maritime Affairs Unit, Via E. Fermi 2749, 21027 Ispra VA, Italy.
