---
title: An_introduction_to_MSE_with_FLR
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document
tags:
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE}
# This chunk set the document environment, so it is hidden
library(knitr)
opts_chunk$set(fig.align='center',
  message=FALSE, warning=FALSE, echo=TRUE, cache=TRUE)
set.seed(1423)
```

Management Strategy Evaluation (MSE) is a framework for evaluating the performance of Harvest Control Rules (HCRs) against prevailing uncertainties. This tutorial introduces the basic steps for building a simple single-species MSE: conditioning the operating model, setting up the observation error model, constructing a simple model-based HCR (based on the ICES MSY approach), performing the MSE simulations (including feedback), and producing performance statistics.

## Required packages

To follow this tutorial you should have installed the following packages:

- CRAN: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/), [FLash](http://www.flr-project.org/FLCore/),[FLXSA](http://www.flr-project.org/FLCore/),[FLBRP](http://www.flr-project.org/FLCore/), [ggplotFL](http://www.flr-project.org/ggplotFL/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("ggplot2"))
install.packages(c("FLa4a","FLash","FLXSA","FLBRP","ggplotFL"), repos="http://flr-project.org/R")
```

```{r, pkgs}
# This chunk loads all necessary packages, trims pkg messages
library(FLa4a)
library(FLash)
library(FLXSA)
library(FLBRP)
library(ggplotFL)
```

# CONDITIONING THE OPERATING MODEL

**Read in stock assessment data**
```{r, dat}
data(ple4)
data(ple4.index)
stk <- ple4; rm("ple4")
idx <- FLIndices(idx=ple4.index); rm("ple4.index")
```
**Set up the iteration and projection window parameters**
```{r, parproj}
it <- 20 # iterations
y0 <- range(stk)["minyear"] # initial data year
dy <- range(stk)["maxyear"] # final data year
iy <- dy+1  # initial year of projection (also intermediate)
fy <- dy+12 # final year
ny <- fy - iy + 1 # number of years to project from intial year
nsqy <- 3 # number of years to compute status quo metrics
```
**Fit stock assessment model a4a**
```{r, a4a}
qmod <- list(~s(age, k=6))
fmod <- ~ te(replace(age, age>9,9), year, k=c(6,8))
mcmc <- 2000
mcsave <- mcmc / it  #this needs to be an integer value
fit <- a4aSCA(stk, idx, fmodel=fmod, qmodel=qmod, fit="MCMC", mcmc = SCAMCMC(mcmc = mcmc, mcsave = mcsave, mcprobe = 0.4))
stk <- stk + fit
stk0 <- qapply(stk, iterMedians) #reduce to keep one iteration only
```
**Fit stock-recruit model**
```{r, sr, message=FALSE}
srbh <- fmle(as.FLSR(stk, model="bevholt"), method="L-BFGS-B", lower=c(1e-6, 1e-6), upper=c(max(rec(stk)) * 3, Inf))
srbh0 <- fmle(as.FLSR(stk0, model="bevholt"), method="L-BFGS-B", lower=c(1e-6, 1e-6), upper=c(max(rec(stk)) * 3, Inf))
srbh.res <- rnorm(it, FLQuant(0, dimnames=list(year=iy:fy)), c(apply(residuals(srbh), 6, mad)))
```
**Calculate reference points and set up the operating model for the projection window**
```{r, refpts}
brp <- brp(FLBRP(stk0, srbh0))
Fmsy <- c(refpts(brp)["msy","harvest"])
Bpa <- 0.5*c(refpts(brp)["msy","ssb"])
stk <- stf(stk, fy-dy, nsqy, nsqy)
```

# SET UP OBSERVATION ERROR MODEL ELEMENTS

**Estimate the indices catchability from the a4a fit (without simulation)**
```{r, idx}
idcs <- FLIndices()
for (i in 1:length(idx)){
    lst <- mcf(list(idx[[i]]@index, stock.n(stk0)))
    idx.lq <- log(lst[[1]]/lst[[2]]) # log catchability of index
    idx.qmu <- idx.qsig <- stock.n(iter(stk,1)) # empty quant
    idx.qmu[] <- yearMeans(idx.lq) # Every year has the same mean catchability
    idx.qsig[] <- log((sqrt(yearVars(idx.lq))/yearMeans(idx.lq))^2 + 1) # Every year has same sd
    idx.q <- FLQuant(NA, dimnames=dimnames(stock.n(stk)))
    idx.q[,ac(dimnames(stock.n(stk))$year[1]:dy)] <- propagate(exp(idx.lq[,ac(dimnames(stock.n(stk))$year[1]:dy)]), it)
    idx.q <- rlnorm(it, idx.qmu, idx.qsig) # Build FLQ of index catchability based on lognormal distribution with mean and sd calculated above
    idx_temp <- idx.q * stock.n(stk)
    idx_temp <- FLIndex(index=idx_temp, index.q=idx.q) # generate initial index
    range(idx_temp)[c("startf", "endf")] <- c(0, 0)
    idcs[[i]] <- idx_temp
}
names(idcs) <- names(idx)
idx<-idcs[1]
```

# SET UP MSE LOOP

## Needed Functions  

**Observation error model**
```{r, oem}
o <- function(stk, idx, assessmentYear, dataYears) {
	# dataYears is a position vector, not the years themselves
	stk0 <- stk[, dataYears]
	# add small amount to avoid zeros
	catch.n(stk0) <- catch.n(stk0) + 0.1
	# Generate the indices - Just data years
	idx0 <- lapply(idx, function(x) x[,dataYears])
  # Generate objserved index
	for (i in 1:length(idx)) index(idx[[i]])[, assessmentYear] <- stock.n(stk)[, assessmentYear]*index.q(idx[[i]])[, assessmentYear]
  list(stk=stk0, idx=idx0, idx.om=idx)
}
```
**XSA assessment model**
```{r, xsa}
xsa <- function(stk0, idx0){
  # Use default XSA settings
    control  <- FLXSA.control(tol = 1e-09, maxit=99, min.nse=0.3, fse=2.0,
                              rage = -1, qage = stk0@range["max"]-1, shk.n = TRUE, shk.f = TRUE,
                              shk.yrs = 5, shk.ages= 5, window = 100, tsrange = 99, tspower = 0)
  # Fit XSA
  fit0 <- FLXSA(stk0, idx0, control)
  # convergence diagnostic (quick and dirty)
  maxit <- c("maxit" = fit0@control@maxit)
  # Update stk0
  stk0   <- transform(stk0, harvest = fit0@harvest, stock.n = fit0@stock.n)
  return(list(stk0 = stk0, converge = maxit))
}
```
**Control object for projections**
```{r, ctrlproj}
getCtrl <- function(values, quantity, years, it){
	dnms <- list(iter=1:it, year=years, c("min", "val", "max"))
	arr0 <- array(NA, dimnames=dnms, dim=unlist(lapply(dnms, length)))
	arr0[,,"val"] <- unlist(values)
	arr0 <- aperm(arr0, c(2,3,1))
	ctrl <- fwdControl(data.frame(year=years, quantity=quantity, val=NA))
	ctrl@trgtArray <- arr0
	ctrl
}
```
## MSE initialisation
```{r, mseinit}
vy <- ac(iy:fy)
TAC <- FLQuant(NA, dimnames=list(TAC="all", year=c(dy,vy), iter=1:it))
TAC[,ac(dy)] <- catch(stk)[,ac(dy)]
TAC[,ac(iy)] <- TAC[,ac(dy)] #assume same TAC in the first intermediate year
ctrl <- getCtrl(c(TAC[,ac(iy)]), "catch", iy, it)
stk <- fwd(stk, control=ctrl, sr=srbh, sr.residuals = srbh.res, sr.residuals.mult = FALSE)
```
## Start the MSE loop
```{r, mseloop}
for(i in vy[-length(vy)]){
  # set up simulations parameters
	ay <- an(i)
	cat(i, " > ")
	vy0 <- 1:(ay-y0) # data years (positions vector) - one less than current year
	sqy <- (ay-y0-nsqy+1):(ay-y0) # status quo years (positions vector) - one less than current year

  # apply observation error
	oem <- o(stk, idx, i, vy0)
	stk0 <- oem$stk
	idx0 <- oem$idx
	idx <- oem$idx.om

  # perform assessment
  out.assess <- eval(call("xsa", stk0, idx0))
  stk0 <- out.assess$stk0
  
  # apply ICES MSY-like Rule to obtain Ftrgt
  flag <- ssb(stk0)[,ac(ay-1)]<Bpa
  Ftrgt <- ifelse(flag,ssb(stk0)[,ac(ay-1)]*Fmsy/Bpa,Fmsy) 

  # project the perceived stock to get the TAC for ay+1
  fsq0 <- yearMeans(fbar(stk0)[,sqy]) # Use status quo years defined above
  ctrl <- getCtrl(c(fsq0, Ftrgt), "f", c(ay, ay+1), it)
  stk0 <- stf(stk0, 2)
  gmean_rec <- c(exp(yearMeans(log(rec(stk0)))))
  stk0 <- fwd(stk0, control=ctrl, sr=list(model="mean", params = FLPar(gmean_rec,iter=it)))
  TAC[,ac(ay+1)] <- catch(stk0)[,ac(ay+1)]

  # apply the TAC to the operating model stock
  ctrl <- getCtrl(c(TAC[,ac(ay+1)]), "catch", ay+1, it)
  stk <- fwd(stk, control=ctrl,sr=srbh, sr.residuals = srbh.res, sr.residuals.mult = FALSE)
}
```

```{r figA, fig.cap="Figure. Operating model results for applying an ICES MSY-like rule"}
plot(stk)+geom_vline(aes(xintercept=as.numeric(ISOdate(iy,1,1))))
```

# References

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* ggplot2: `r packageVersion('ggplot2')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission Joint Research Centre (JRC), Institute for the Protection and Security of the Citizen (IPSC), Maritime Affairs Unit, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>
