---
title: Reading data into FLR 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document
tags:
license: Creative Commons CC-BY SA
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE}
# This chunk set the document environment, so it is hidden
library(knitr)
opts_chunk$set(fig.align='center',
  message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE)
options(width=50)
set.seed(1423)
```

PD 
This tutorial details methods for reading data and generating the FLStock() object class.

## Required packages


Youen !!!

To follow this tutorial you should have installed the following packages:

- CRAN: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/), [ggplotFL](http://www.flr-project.org/ggplotFL/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("ggplot2"))
install.packages(c("ggplotFL"), repos="http://flr-project.org/R")
```

```{r, pkgs}
# This chunk loads all necessary packages, trims pkg messages
library(FLCore)
library(ggplotFL)
```

# SECTION

PD


```{r figA}
# This is an example chunk for a figure
plot(FLQuant(rnorm(200), dim=c(2,20)))
```

## SubSECTION : Reading files (csv, dat, ...) - YV

## Subsection : Reading fisheries data formats (VPA etc..) - PD

## SUbsection : Reshaping data as a matrix - YV

## Subection :  Making an FLQuant object - PD

## Subsection : Description, units, ranges etc.. - PD

# References


# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* ggplot2: `r packageVersion('ggplot2')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission Joint Research Centre (JRC), Institute for the Protection and Security of the Citizen (IPSC), Maritime Affairs Unit, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>
