---
title: Reading data into FLR 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document
tags:
license: Creative Commons CC-BY SA
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE}
# This chunk set the document environment, so it is hidden
library(knitr)
opts_chunk$set(fig.align='center',
  message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE)
options(width=50)
set.seed(1423)
```

This tutorial details methods for reading various formats of data into R for generating the FLStock() object class.

## Required packages


Youen !!!

To follow this tutorial you should have installed the following packages:

- CRAN: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/), [ggplotFL](http://www.flr-project.org/ggplotFL/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("ggplot2"))
install.packages(c("ggplotFL"), repos="http://flr-project.org/R")
```

```{r, pkgs}
# This chunk loads all necessary packages, trims pkg messages
library(FLCore)
library(ggplotFL)
```

# SECTION

PD

## SubSECTION : Reading files (csv, dat, ...) - YV

Fisheries data are generally stored in different format (cvs, excel, SAS...). R provides tools to read and import data from simple text files to more advanced SAS files or databases.
https://www.datacamp.com/community/tutorials/importing-data-r-part-two#gs.kNzBd5k is a nice tutorial to quickly import data into R.

Your data are stored in a folder in your computer or a server. You have to tell R what is the path to the data.
You can check the working directory already active in your R session using the command getwd(). 
To set the working directory use setwd("directory name"). Case is important, use // or \ for separating folders and directories in Windows.


This tutorial will give some exemples but regardless the format, the different steps are:
- Finding the right function to import data into R
- Reshaping the data as a matix 
- creating an FLQuant objetc

### importing CSV files

There is many ways of reading csv files. 
read.table with 'header', 'sep', 'dec' and 'row.names' options will allow you reading all .csv and .txt files

The 

``` {r }
# Read this in using read.table() with default options

catch.n <- read.csv("src/Data/catch_numbers.csv",row=1)

# We have read in the data as a data.frame
class(catch.n)
# There is an FLQuant contructor that uses a data.frame, but here our data.frame is not set up the right way
# Instead we can convert the object to a matrix
catch.n.matrix <- as.matrix(catch.n)
catch.n.matrix
# We need to specify the dimnames
catch.n.flq <- FLQuant(catch.n.matrix, dimnames=list(age=1:7, year = 1957:2011))
catch.n.flq

```

## Reading common fisheries data formats 

FLCore contains functions for reading in fish stock data in common pre-defined formats. To read a single variable (e.g. numbers-at-age, maturity-at-age) from the **Lowestoft VPA** format you use the `readVPA` function. The following example reads the catch numbers-at-age for herring:

```{r, readVPA}
catch.n <- readVPAFile(file.path('src','Data','her-irlw',"canum.txt"))
class(catch.n)
```
This can be repeated for each of the data files. 

Alternatively, if you have the full information for a stock in the **Lowestoft VPA**, **Adapt**, **CSA** or **ICA** format you can read in together using the `readFLStock` function. Here, you point the function to the index file, with all other files in the same directory:

```{r, readFLStock}
her <- readFLStock(file.path('src','Data','her-irlw','index.txt'))
class(her)
```
Which we can see correctly formats the data as an `FLStock` object.

```{r, readFLStock2}
summary(her)
```
Note that the units for the slots have not been set. We will deal with this in the section below.

In addition, this object only contains the input data for the stock assessment, not any estimated values (e.g. harvest rates, stock abundances). You can add these to the object as follows:

```{r, Add values}
her@stock.n <- readVPAFile(file.path('src','Data','her-irlw',"n.txt"))
```

## SUbsection : Reshaping data as a matrix - YV

## Subection :  Making an FLQuant object - PD

## Subsection : Description, units, ranges etc.. - PD

# References


# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* ggplot2: `r packageVersion('ggplot2')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission Joint Research Centre (JRC), Institute for the Protection and Security of the Citizen (IPSC), Maritime Affairs Unit, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>
