---
title: Reading data into FLR 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  github_document
tags:
license: Creative Commons CC-BY SA
---
 
```{r, ini, echo=FALSE, results='hide', message=FALSE}
# This chunk set the document environment, so it is hidden
library(knitr)
opts_chunk$set(fig.align='center',
  message=FALSE, warning=FALSE, echo=TRUE, cache=FALSE)
options(width=50)
set.seed(1423)
```

This tutorial details methods for reading various formats of data into R for generating the `FLStock` object class.

## Required packages

To follow this tutorial you should have installed the following packages:

- CRAN: [ggplot2](https://cran.r-project.org/web/packages/ggplot2/index.html)
- FLR: [FLCore](http://www.flr-project.org/FLCore/); [FLFleet](http://www.flr-project.org/FLFleet/); [ggplotFL](http://www.flr-project.org/ggplotFL/)

You can do so as follows,

```{r, eval=FALSE}
install.packages(c("ggplot2"))
install.packages(c("ggplotFL"), repos="http://flr-project.org/R")
```

```{r, pkgs}
# This chunk loads all necessary packages, trims pkg messages
library(FLCore); library(FLFleet)
library(ggplotFL)
```

# FLStock objects 

This section covers methods for reading in the data required to construct `FLStock` objects.

## Reading files (csv, dat, ...)

Fisheries data are generally stored in different format (cvs, excel, SAS...). R provides tools to read and import data from simple text files to more advanced SAS files or databases. [Datacamp](https://www.datacamp.com/community/tutorials/importing-data-r-part-two#gs.kNzBd5k) is a nice tutorial to quickly import data into R.

Your data are stored in a folder in your computer or a server. You have to tell R what is the path to the data.
You can check the working directory already active in your R session using the command getwd(). 
To set the working directory use setwd("directory name"). Case is important, use // or \ for separating folders and directories in Windows.

This tutorial will give some examples but regardless the format, the different steps are:
- Finding the right function to import data into R
- Reshaping the data as a matrix 
- creating an FLQuant object

### Importing files into R (exemple of csv file)

There is many ways of reading csv files. 
`read.table` with 'header', 'sep', 'dec' and 'row.names' options will allow you reading all .csv and .txt files

The `read.csv` or `read.csv2` function are very useful to read csv files. 

``` {r }
catch.n <- read.csv("src/Data/catch_numbers.csv",row=1)
```

We have read in the data as a data.frame

``` {r }
class(catch.n)
```

The data are now in your R environment, before creating a **FLQuant** object, you need to make sure it is consistent with the type of object and formatting that is needed to run the `FLQuant()` function. To get information on the structure and format needed type ?FLQuant in your R Console.


## Reshaping data as a matrix 

FLQuant accept 'vector', 'array' or 'matrix'. 
We can convert the object catch.n to a matrix

``` {r }
catch.n.matrix <- as.matrix(catch.n)
catch.n.matrix[,1:8]
```

A `FLQuant` object is made of six dimensions. The name of the first dimension can be altered by the user from its default, quant. This could typically be age or length for data related to natural populations. The only name not accepted is 'cohort', as data structured along cohort should be stored using the `FLCohort` class instead. Other dimensions are always names as follows: year, for the calendar year of the data point; unit, for any kind of division of the population, e.g. by sex; season, for any temporal strata shorter than year; area, for any kind of spatial stratification; and iter, for replicates obtained through bootstrap, simulation or Bayesian analysis.

When importing catch number for example, the input object needs to be formatted as such: age or length in the first dimension and years in the second dimension. If the object is not formatted in the right way, you can use the `reshape()` function from the package [reshape2](https://cran.r-project.org/web/packages/reshape2/index.html).


## Making an FLQuant object 

We need to specify the dimnames
``` {r }
catch.n.flq <- FLQuant(catch.n.matrix, dimnames=list(age=1:7, year = 1957:2011))
catch.n.flq[,1:7]

```

## Reading common fisheries data formats 

FLCore contains functions for reading in fish stock data in commonly used formats. To read a single variable (e.g. numbers-at-age, maturity-at-age) from the **Lowestoft VPA** format you use the `readVPA` function. The following example reads the catch numbers-at-age for herring:

```{r, readVPA}
# Read from a VPA text file
catch.n <- readVPAFile(file.path('src','Data','her-irlw',"canum.txt"))
class(catch.n)
```
This can be repeated for each of the data files. In addition, functions are available for Multifan-CL format  `readMFCL` and ADMB `readADMB`.

Alternatively, if you have the full information for a stock in the **Lowestoft VPA**, **Adapt**, **CSA** or **ICA** format you can read in together using the `readFLStock` function. Here, you point the function to the index file, with all other files in the same directory:

```{r, readFLStock}
# Read a collection of VPA files, pointing to the Index file:
her <- readFLStock(file.path('src','Data','her-irlw','index.txt'))
class(her)
```
Which we can see correctly formats the data as an `FLStock` object.

```{r, readFLStock2}
summary(her)
```
Note: the units for the slots have not been set. We will deal with this in the next section.

In addition, this object only contains the input data for the stock assessment, not any estimated values (e.g. harvest rates, stock abundances). You can add these to the object as follows:

```{r, AddMissingAssessmentData}
her@stock.n <- readVPAFile(file.path('src','Data','her-irlw',"n.txt"))
print(her@stock.n[,ac(2007:2011)]) # only print 2007:2011

her@harvest <- readVPAFile(file.path('src','Data','her-irlw',"f.txt"))

```
Now we have a fully filled `FLStock` object. But let's check the data are consistent.

```{r, CheckConsistency}
# The sum of products (SOP)
apply(her@landings.n * her@landings.wt, 2, sum)[,ac(2007:2011)]
# and the value read in from the VPA file
her@landings[,ac(2007:2011)]

## They are not the same!!  We correct the landings to be the same as the SOP - there is a handy function for this purpose
her@landings <- computeLandings(her)

# In addition, there is no discard information
her@discards.wt[,ac(2005:2011)]
her@discards.n[,ac(2005:2011)]
# Set up the discards and catches
her@discards.wt   <- her@landings.wt
her@discards.n[]  <- 0
her@discards      <- computeDiscards(her)
her@catch         <- her@landings
her@catch.wt      <- her@landings.wt
her@catch.n       <- her@landings.n

```
Functions are available to `computeLandings`, `computeDiscards`, `computeCatch` and `computeStock`. These functions take the argument `slot = 'catch'`, `slot = 'wt'` and `slot = 'n'` to compute the total weight, individual weight and numbers respectively, in addition to `slot = 'all'`. 

## Description, units, ranges etc..

Before we are finished, we want to ensure the units and range references are correct. This is important as the derived calculations require the correct scaling.

First, let's ensure an appropriate name and description are assigned:

```{r, Descriptions}
summary(her)

#name and descriptions
her@name # ok
her@desc # ok

# Set the Fbar range for the stock 
her@range[c('minfbar','maxfbar')] # ok, but can be filled with c(min,max)

# set the plus group
her@range['plusgroup']  <- 7 # final year is a plusgroup

## Units
units(her@catch)    <- units(her@discards)    <- units(her@landings)    <- units(her@stock)    <- 'tonnes'
units(her@catch.n)  <- units(her@discards.n)  <- units(her@landings.n)  <- units(her@stock.n)  <- '1000'
units(her@catch.wt) <- units(her@discards.wt) <- units(her@landings.wt) <- units(her@stock.wt) <- 'kg'
units(her@harvest) <- 'f'

```
This should now have the correct units defined:

```{r, Plot}
summary(her)
plot(her) + theme_bw() # using the simple bw theme
```

# FLIndices

Two solutions can be used to read abundance indices into FLR. 

## Reading from common fisheries data formats 

If your data are formated in a **Lowestoft VPA** format then FLCore contains functions for reading in indices. To read an abundance index, you use the `readFLIndices` function. The following example reads the index from `ple4` exemple:

```{r }
indices <- readFLIndices('src/Data/ple4_ISIS.txt')
```

Using this function, `slot indices@names` is already filled by `r indices@names`, and the information `slot indices@range` too.

## Reading from flat files 

If your data are not formated in a **Lowestoft VPA** format, then you and read them using `read.table` for exemple.

```{r }
indices <- read.table('src/Data/ple4Index1.txt')
```
which needs to be transformed in FLQuant
```{r }
indices <- FLQuant(as.matrix(indices), dimnames=list(age=1:8, year = 1985:2008))
```
And in FLIndex
```{r }
indices <- FLIndex(indices)
```
And then in FLIndices
```{r }
indices <- FLIndices(indices)
```

`slot indices@range` needs to be filled in with the end and start date of the tunning series

```{r }
indices[[1]]@range[c('startf', 'endf')] <- c(0.66,0.75)
```




# FLFleets

Reading data on fleets into an `FLFleet` object is complicated by the multi-layer structure of the object. The object is defined so that:

```{r, level table, echo = F}
kable(data.frame(Level = c(1,2,3),
		 Class = c('FLFleet','FLMetier(s)','FLCatch(es)'),
		 Contains = c('variables relating to vessel level activity',
			     'variables relating to fishing level activity',
			     'variables relating to stock catches')))
```

Here are the slots for each level:


```{r, FLFleet slots}
# FLFleet level
summary(FLFleet())
# FLMetier level
summary(FLMetier())
# FLCatch level
summary(FLCatch())
```

Due to the different levels, units and dimensions of the variables and the potentially high number of combinations of fleets, métier and stocks in a mixed fishery - getting the full data into an FLFleets object can be an onerous task.

A way of simplifying the generation of the fleet object is to ensure all the data are in a csv file with the following structure:

```{r, Fleetdata}
kable(data.frame(Fleet = c('Fleet1', 'Fleet2'),
		 Metier = c('Metier1', 'Metier1'),
		 Stock = c('Stock1', 'Stock2'),
		 type = c('landings.n', 'landings.wt'),
		 age = c(1,1),
		 year = c(2011,2011),
		 unit = c(1,1),
		 season = c('all', 'all'),
		 area = c('unique', 'unique'),
		 iter = c(1,1),
		 data = c(254,0.3)))
```


To generate the required structure, you can then read in the file and generate the object using an `lapply` function:

```{r, Generating FLFleets, eval = F}
# Example of generating fleets
fl.nam <- unique(data$Fleet) # each of the fleets

yr.range <- 2005:2011 # year range of the data - must be same, even if filled with NAs or 0s

# empty FLQuant for filling with right dimensions
fq  <- FLQuant(dimnames = list(year = yr.range), quant = 'age')

### Fleet level slots ###
fleets <- FLFleet(lapply(fl.nam, function(Fl) {

# blank quants with the same dims
eff <- cap <- crw <- cos.fl <- fq

# fleet effort
eff[,ac(yr.range)] <- data$data[data$Fleet == Fl & data$type == 'effort']
units(eff) <- '000 kw days'

## Repeat for each fleet level variables (not shown) ##

### Metier level slots ###
met.nam  <- unique(data$Metier[data$Fleet == Fl]) # metiers for fleet
met.nam  <- met.nam[!is.na(met.nam)] # exclude the fleet level data

metiers  <- FLMetiers(lapply(met.nam, function(met) {

# blank quants
effmet <- cos.met <- fq

# effort share for metier
effmet[,ac(yr.range)] <- data$data[data$Fleet == Fl & data$Metier & data$type == 'effshare']
units(effmet)  <- NA

## Repeat for each metier level variables (not shown) ##


sp.nam <- unique(data$stock[data$Fleet == Fl & data$Metier == met]) # stocks caught by metier
sp.nam <- sp.nam[!is.na(sp.nam)] # exclude fleet and metier level data

catch <- FLCatches(lapply(sp.nam, function(S){
print(S)

# Quant dims may be specific per stock
la.age <- FLQuant(dimnames = list(age = 1:7, year = yr.range, quant = 'age'))
la.age[,ac(yr.range)] <- data$data[data$Fleet == Fl & data$Metier == met & data$Stock == S & data$type == 'landings.n']
units(la.age) <- '1000'

## Repeat for all stock level variables (not shown) ##

# Build F
res <- FLCatch(range = yr.range, name = S, landings.n = la.age,...)

## Compute any missing slots, e.g.
res@landings <- computeLandings(res)

return(res) # return filled FLCatch

})) # End of FLCatches

# Fill an FLMetier with all the stock catches
m <- FLMetier(catches = catch, name = met)
m@effshare  <- effmet
m@vcost <- vcost

		 })) # end of FLMetiers

fl <- FLFleet(metiers = metiers, name = Fl, effort = ef,...) # fill with all variables
return(fl)

		 }))

names(fleets) <- fl.nam

```

You should now have a multilevel object with `FLFleets`, `FLMetiers` and `FLCatches`.


# References

None

# More information

* You can submit bug reports, questions or suggestions on this tutorial at <https://github.com/flr/doc/issues>.
* Or send a pull request to <https://github.com/flr/doc/>
* For more information on the FLR Project for Quantitative Fisheries Science in R, visit the FLR webpage, <http://flr-project.org>.

## Software Versions

* `r version$version.string`
* FLCore: `r packageVersion('FLCore')`
* ggplotFL: `r packageVersion('ggplotFL')`
* ggplot2: `r packageVersion('ggplot2')`
* **Compiled**: `r date()`

## License

This document is licensed under the [Creative Commons Attribution-ShareAlike 4.0 International](https://creativecommons.org/licenses/by-sa/4.0) license.

## Author information

**Iago MOSQUEIRA**. European Commission Joint Research Centre (JRC), Institute for the Protection and Security of the Citizen (IPSC), Maritime Affairs Unit, Via E. Fermi 2749, 21027 Ispra VA, Italy. <https://ec.europa.eu/jrc/>


